
```{r}
library(papaja)
```


```{r, echo = F}
# Data
here::i_am("data/polish/data_polish.R")
source("data/polish/data_polish.R")
```
```{r}
# societal languages of the multilingual children

write.csv(by_child_pl, "by_child_pl.csv")

# dzieci_kraje_jezyki_wide.csv
# country_raw	country	kod_pocztowy	country_code


```


The Polish validation study included `r sum(by_child_pl$lang_group == "multilingual")` data from bi- and multilingual families. Altogether, for 28 children, parents reported that the children had regular contact with languages other than Polish. Of these, 25 parents specified the languages involved, while 3 did not. Eighteen children were bilingual: 12 had regular contact with English, 5 with Norwegian, and 1 with Russian. Seven children were multilingual. All seven had regular contact with Polish and English; additionally, 3 had contact with Italian, 2 with Spanish, 1 with Turkish, and 1 with Norwegian.

Though the bi-/multilingual group is small in size, we decided to explore their correlation coefficients (non-parametric Spearman's rho) and found these were similar to those found for Polish monolingual children (see Table \@ref(tab:corr-by-langgroup)). Since the correlation coefficients were similar in both groups, and as the bi- and multilingual group was was too small to warrant a separate analysis, we decided to include them in the Polish dataset alongside the Polish monolinguals.

```{r corr-by-langgroup, include = T}

# TODO: 
# 1. the table could use some re-styling (lgs as columns?)
# 2. move to Supplementary (separate .Rmd?)

# format p-values: "< .001" for very small ones, otherwise 3 decimals
fmt_p <- function(p) ifelse(is.na(p), NA,
                            ifelse(p < .001, "< .001", sprintf("%.3f", p)))

sig <- function(x, is_sig) ifelse(is_sig, paste0("\\textbf{", x, "}"), x)
# or colour: paste0("\\textcolor{blue}{", x, "}")

spearman_by_lang <- function(data, x, y, label) {
  data %>%
    group_by(lang_group) %>%
    summarise(
      rho = cor({{ x }}, {{ y }}, method = "spearman", use = "complete.obs"),
      n_complete = sum(complete.cases({{ x }}, {{ y }})),
      p = if (n_complete > 2)
            cor.test({{ x }}, {{ y }}, method = "spearman", exact = FALSE)$p.value
          else NA_real_,
      ci_low  = spearman_ci(rho, n_complete)[1],
      ci_high = spearman_ci(rho, n_complete)[2],
      n = n(),
      .groups = "drop") %>%
    select(-n_complete) %>%
    mutate(correlation = label)
}

lang_cor_table_raw <- bind_rows(
  spearman_by_lang(by_child_pl, cat_theta,  score_full, "Ability from CDI-CAT $\\sim$ full CDI score"),
  spearman_by_lang(by_child_pl, cat_theta,  full_theta, "Ability from CDI-CAT $\\sim$ ability from full CDI"),
  spearman_by_lang(by_child_pl, full_theta, score_full, "Ability from full CDI $\\sim$ full CDI score"))

lang_cor_table <- lang_cor_table_raw %>%
  mutate(lang_group = gsub("_", " ", lang_group),
         is_sig = !is.na(p) & p < .05,
         ci     = ifelse(is.na(ci_low), NA,
                         sprintf("[%.2f, %.2f]", ci_low, ci_high)),
         rho    = sprintf("%.2f", rho),
         p      = fmt_p(p),
         across(everything(), ~ sig(as.character(.x), is_sig))) %>%
  select(correlation, lang_group, rho, ci, p, n) %>%
  rename("Correlation" = "correlation", "Language group" = "lang_group",
         "$\\rho$" = "rho", "95\\% CI" = "ci", "$p$" = "p")

apa_table(lang_cor_table, 
          escape = FALSE,
          align = c("l", "l", "c", "c", "c", "c"),
          font_size = "small",
          note = "Significant correlations ($p$ < .05) are shown in bold.",
          caption = "Spearman's correlations for monolingual and multilingual children in the Polish dataset.",
          placement = "H")

```
