
```{r, echo = F}
# Data
here::i_am("data/polish/data_polish.R")
source("data/polish/data_polish.R")
```
```{r}
# societal languages of the multilingual children

write.csv(by_child_pl, "by_child_pl.csv")

# dzieci_kraje_jezyki_wide.csv
# country_raw	country	kod_pocztowy	country_code


```


The Polish validation study included `r sum(by_child_pl$lang_group == "multilingual")` data from bi- and multilingual families. Altogether, for 28 children, parents reported that the children had regular contact with languages other than Polish. Of these, 25 parents specified the languages involved, while 3 did not. Eighteen children were bilingual: 12 had regular contact with English, 5 with Norwegian, and 1 with Russian. Seven children were multilingual. All seven had regular contact with Polish and English; additionally, 3 had contact with Italian, 2 with Spanish, 1 with Turkish, and 1 with Norwegian.

Though the bi-/multilingual group is small in size, we decided to explore their correlation coefficients (non-parametric Spearman's rho) and found these were similar to those found for Polish monolingual children (see Table \@ref(tab:corr-by-langgroup)). Since the correlation coefficients were similar in both groups, and as the bi- and multilingual group was was too small to warrant a separate analysis, we decided to include them in the Polish dataset alongside the Polish monolinguals.

```{r corr-by-langgroup, include = T}

# TODO: 
# 1. the table could use some re-styling (lgs as columns?)
# 2. move to Supplementary (separate .Rmd?)

cor1 <- by_child_pl %>%
  group_by(lang_group) %>%
  summarise(r = cor(cat_theta, score_full, method = "spearman"), n = n()) %>%
  mutate(correlation = "Ability from CDI-CAT ~ full CDI score")

cor2 <- by_child_pl %>%
  group_by(lang_group) %>%
  summarise(r = cor(cat_theta, full_theta, method = "spearman"), n = n()) %>%
  mutate(correlation = "Ability from CDI-CAT ~ ability from full CDI")

cor3 <- by_child_pl %>%
  group_by(lang_group) %>%
  summarise(r = cor(full_theta, score_full, method = "spearman"), n = n()) %>%
  mutate(correlation = "Ability from full CDI ~ full CDI score")

lang_cor_table <- bind_rows(cor1, cor2, cor3)
rm(cor1, cor2, cor3)

apa_table(lang_cor_table, caption = "Spearman's correlations for monolingual and multilingual children in the Polish dataset", placement = "H")

```
