---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "Karolina Muszyńska"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "University of Warsaw, Faculty of Psychology, ul. Stefana Banacha 2D, 02-097 Warsaw, Poland"
    email         : "karolina.muszynska@psych.uw.edu.pl"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "XXX"
    affiliation   : "2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "University of Warsaw, Faculty of Psychology"
  - id            : "2"
    institution   : "Stanford University"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")

library(tidyverse)
library(Multilada)
library(here)


```

```{r analysis-preferences}
# Seed for random number generation
set.seed(123)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = F, include = F, warning = F, message = F, out.width="60%")
```


```{r}
# Data
source(here("data/polish/data_polish.R"))
source(here("data/english/data_ame_english.R"))
```

# Methods
(to paste from [google doc](https://docs.google.com/document/d/1MPInUTyY2Y81EtiiK0tUOo8vG3zowfqiXjpBF0p8Mzs/edit?tab=t.0))

## Participants
(to paste from [google doc](https://docs.google.com/document/d/1MPInUTyY2Y81EtiiK0tUOo8vG3zowfqiXjpBF0p8Mzs/edit?tab=t.0))
Mention the bilinguals and multilinguals

## Material
(to paste from [google doc](https://docs.google.com/document/d/1MPInUTyY2Y81EtiiK0tUOo8vG3zowfqiXjpBF0p8Mzs/edit?tab=t.0))

## Procedure
(to paste from [google doc](https://docs.google.com/document/d/1MPInUTyY2Y81EtiiK0tUOo8vG3zowfqiXjpBF0p8Mzs/edit?tab=t.0))

## Data analysis


# Results

## Psychometric properties of the two CAT-CDIs

Our first aim was to examine whether CAT-CDIs in American English and Polish demonstrate comparable psychometric properties. To that end, we revisit the psychometric properties reported for the American English CAT-CDI (word production) in Kachergis et al. (2022) and compare those to the data from Polish CAT-CDI (Words and Sentences).

<!-- AmE r values copied from Kachergis et al. 2022 -->
```{r}
cor(by_child_pl$cat_theta, by_child_pl$score_full) # AmE:.86, PL: 0.86
cor(by_child_pl$cat_theta, by_child_pl$full_theta) # AmE: .92, PL: .92
cor(by_child_pl$full_theta, by_child_pl$score_full) # AmE: .95, PL: 0.94
```

We found similarly strong correlations in the two languages between the abilities estimated from CDI-CAT and full CDI scores (American English and Polish: $r$ = .86), the abilities estimated from the CDI-CAT and abilities estimated from full CDI (American English and Polish: $r$ = .92), and the abilities estimated from the full CDI and the full CDI scores (American English: $r$ = .95, Polish: $r$ = `r cor(by_child_pl$full_theta, by_child_pl$score_full)`). The abilities estimated from the CDI-CAT and the full CDI scores were also strongly correlated within individual age groups (see Table \@ref(tab:corr-by-age)).


```{r corr-by-age, include = T}
# Correlations by age
# to-do: merge both tables?
# English
valid_tab <- prod_s %>% 
  mutate(age_group = cut(age_full, breaks=seq(12,36,3), right=F, include.lowest = T)) %>%
  group_by(age_group) %>%
  summarise(r = cor(fullTheta, catTheta), n=n()) #%>% kableExtra::kable(digits=2)

valid_tab_wide = rbind(round(valid_tab$r, 2), as.character(valid_tab$n))
colnames(valid_tab_wide) = as.character(valid_tab$age_group)
row.names(valid_tab_wide) = c('r ability CAT vs full CDI', 'N')

apa_table(valid_tab_wide, caption="American English: Correlations between ability estimated by CAT-CDI and ability estimated from full CDI by children's age") 

# Polish
age_cor_table <- by_child_pl %>% 
  mutate(age_group = cut(age_CAT, breaks=seq(12,36,3), right=F, include.lowest = T)) %>%
  group_by(age_group) %>%
  summarise(r = cor(full_theta, cat_theta), n=n())

age_cor_table_wide = rbind(round(age_cor_table$r, 2), as.character(age_cor_table$n))
colnames(age_cor_table_wide) = as.character(age_cor_table$age_group)
row.names(age_cor_table_wide) = c('r ability CAT vs full CDI', 'N')

apa_table(age_cor_table_wide, caption = "Polish: Correlations between ability estimated by CAT-CDI and ability estimated from full CDI by children's age")
```


The Polish validation study included `r sum(by_child_pl$lang_group == "multilingual")` data from bi- and multilingual families. Though it is a small group, we decided to explore their correlation coefficients (non-parametric Spearman's rho) and found these were similar to those found for Polish monolingual children (see Table X in Supplementary Materials). <!-- TO DO -->

```{r, include = T}

cor1 <- by_child_pl %>%
  group_by(lang_group) %>%
  summarise(r = cor(cat_theta, score_full, method = "spearman"), n = n()) %>%
  mutate(correlation = "Ability from CDI-CAT ~ full CDI score")

cor2 <- by_child_pl %>%
  group_by(lang_group) %>%
  summarise(r = cor(cat_theta, full_theta, method = "spearman"), n = n()) %>%
  mutate(correlation = "Ability from CDI-CAT ~ ability from full CDI")

cor3 <- by_child_pl %>%
  group_by(lang_group) %>%
  summarise(r = cor(full_theta, score_full, method = "spearman"), n = n()) %>%
  mutate(correlation = "Ability from full CDI ~ full CDI score")

lang_cor_table <- bind_rows(cor1, cor2, cor3)
rm(cor1, cor2, cor3)

apa_table(lang_cor_table, caption = "Supplementary Material: Table S1 - Spearman's correlations for monolingual and multilingual children in the Polish dataset")

```






# Discussion


\newpage

# References

::: {#refs custom-style="Bibliography"}
:::

`r cite_r("r-references.bib")`
