---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "Karolina Muszyńska"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "University of Warsaw, Faculty of Psychology, ul. Stefana Banacha 2D, 02-097 Warsaw, Poland"
    email         : "karolina.muszynska@psych.uw.edu.pl"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "To fill in later on"
    affiliation   : "2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "University of Warsaw, Faculty of Psychology"
  - id            : "2"
    institution   : "To fill in later on"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}

library(papaja)
r_refs("r-references.bib")

library(tidyverse)
library(Multilada)
library(here)


```

```{r analysis-preferences}
# Seed for random number generation
set.seed(123)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = F, include = F, warning = F, message = F, out.width="100%")
```


```{r, echo = F}
# Data
source(here("data/polish/data_polish.R"))
source(here("data/english/data_ame_english.R"))
```

# Methods

The data was gathered with CAT-CDIs in American English and Polish. The American English CAT-CDI was created from the items that were common to three CDIs: Words and Gestures, Words and Sentences and CDI-III. With these items, two CAT-CDIs were created, one for word production and another for word comprehension (see Kachergis et al. 2022 for more details). The Polish CAT-CDIs were created following a largely similar procedure as the ones in American English, but there was a separate CDI-CAT developed for each CDI version (CDI: Words and Gestures and CDI: Words and Sentences) and for each CDI:WG subscale (word production, word comprehension, gesture use) (Krajewski et al. in preparation). Here we show data from the validation study of American English CAT-CDI word production (i.e. 679 items) and Polish CAT-CDI version of WS: Words and Sentences (i.e. 666 items). There are 395 words that appear in both CDI-CAT language versions. 


## Participants

In the American English validation study, the participants were 204 parents of children aged 15 to 36 months (*M* = `r round(mean(kept_demo$age_cat),2)`, *Mdn* = `r round(median(kept_demo$age_cat),2)`, SD = `r round(sd(kept_demo$age_cat),2)`). There were 103 girls in the sample (`r round(sum(kept_demo$sex_cat == "F")/count(kept_demo)*100,2)`%). The sample was ethnically diverse, including White, Black, Latino/Hispanic, Asian, Native American, and other races/ethnicities. The average
number of years of education attained by the primary caregiver was `r round(mean(kept_demo$primary_education_cat),2)` (SD = `r round(sd(kept_demo$primary_education_cat),2)`).

In the Polish validation study, the participants were 113 parents of children aged 18 to 34 months (*M* = `r round(mean(by_child_pl$age_CAT),2)`, *Mdn* = `r round(median(by_child_pl$age_CAT),2)`, SD = `r round(sd(by_child_pl$age_CAT),2)`). There were 49 girls in the sample (`r round(sum(by_child_pl$sex == "dziewczynka")/count(by_child_pl)*100,2)`%). The sample was White (reflecting the limited ethnic diversity within the Polish population). However, there were 28 children multilingual with Polish (as their home language) and some other (majority) language. [todo: add info on parental edu]

```{r}
kept_demo$primary_education_cat
```



## Material
(to paste from [google doc](https://docs.google.com/document/d/1MPInUTyY2Y81EtiiK0tUOo8vG3zowfqiXjpBF0p8Mzs/edit?tab=t.0))

## Procedure
(to paste from [google doc](https://docs.google.com/document/d/1MPInUTyY2Y81EtiiK0tUOo8vG3zowfqiXjpBF0p8Mzs/edit?tab=t.0))

## Data analysis


# Results

## Psychometric properties of the two CAT-CDIs

Our first aim was to examine whether CAT-CDIs in American English and Polish demonstrate comparable psychometric properties. To that end, we revisit the psychometric properties reported for the American English CAT-CDI (word production) in Kachergis et al. (2022) and compare those to the data from Polish CAT-CDI (Words and Sentences).

<!-- AmE r values copied from Kachergis et al. 2022 -->
```{r}
cor(by_child_pl$cat_theta, by_child_pl$score_full) # AmE:.86, PL: 0.86
cor(by_child_pl$cat_theta, by_child_pl$full_theta) # AmE: .92, PL: .92
cor(by_child_pl$full_theta, by_child_pl$score_full) # AmE: .95, PL: 0.94
```

We found similarly strong correlations in the two languages between the abilities estimated from CDI-CAT and full CDI scores (American English and Polish: $r$ = .86), the abilities estimated from the CDI-CAT and abilities estimated from full CDI (American English and Polish: $r$ = .92), and the abilities estimated from the full CDI and the full CDI scores (American English: $r$ = .95, Polish: $r$ = `r cor(by_child_pl$full_theta, by_child_pl$score_full)`). The abilities estimated from the CDI-CAT and the full CDI scores were also strongly correlated within individual age groups (see Table \@ref(tab:corr-by-age)).


```{r corr-by-age, include = T}

# TODO:
# 1. merge both tables?
# 2. can we specify place in text where the tables should be placed (rn it seems random)

# English
valid_tab <- prod_s %>% 
  mutate(age_group = cut(age_full, breaks=seq(12,36,3), right=F, include.lowest = T)) %>%
  group_by(age_group) %>%
  summarise(r = cor(fullTheta, catTheta), n=n()) #%>% kableExtra::kable(digits=2)

valid_tab_wide = rbind(round(valid_tab$r, 2), as.character(valid_tab$n))
colnames(valid_tab_wide) = as.character(valid_tab$age_group)
row.names(valid_tab_wide) = c('r ability CAT vs full CDI', 'N')

apa_table(valid_tab_wide, caption="American English: Correlations between ability estimated by CAT-CDI and ability estimated from full CDI by children's age", placement = "H") 

# Polish
age_cor_table <- by_child_pl %>% 
  mutate(age_group = cut(age_CAT, breaks=seq(12,36,3), right=F, include.lowest = T)) %>%
  group_by(age_group) %>%
  summarise(r = cor(full_theta, cat_theta), n=n())

age_cor_table_wide = rbind(round(age_cor_table$r, 2), as.character(age_cor_table$n))
colnames(age_cor_table_wide) = as.character(age_cor_table$age_group)
row.names(age_cor_table_wide) = c('r ability CAT vs full CDI', 'N')

apa_table(age_cor_table_wide, caption = "Polish: Correlations between ability estimated by CAT-CDI and ability estimated from full CDI by children's age", placement = "H")
```


The Polish validation study included `r sum(by_child_pl$lang_group == "multilingual")` data from bi- and multilingual families. Though it is a small group, we decided to explore their correlation coefficients (non-parametric Spearman's rho) and found these were similar to those found for Polish monolingual children (see Table \@ref(tab:corr-by-langgroup) in Supplementary Materials). 

```{r corr-by-langgroup, include = T}

# TODO: 
# 1. the table could use some re-styling (lgs as columns?)
# 2. move to Supplementary (separate .Rmd?)

cor1 <- by_child_pl %>%
  group_by(lang_group) %>%
  summarise(r = cor(cat_theta, score_full, method = "spearman"), n = n()) %>%
  mutate(correlation = "Ability from CDI-CAT ~ full CDI score")

cor2 <- by_child_pl %>%
  group_by(lang_group) %>%
  summarise(r = cor(cat_theta, full_theta, method = "spearman"), n = n()) %>%
  mutate(correlation = "Ability from CDI-CAT ~ ability from full CDI")

cor3 <- by_child_pl %>%
  group_by(lang_group) %>%
  summarise(r = cor(full_theta, score_full, method = "spearman"), n = n()) %>%
  mutate(correlation = "Ability from full CDI ~ full CDI score")

lang_cor_table <- bind_rows(cor1, cor2, cor3)
rm(cor1, cor2, cor3)

apa_table(lang_cor_table, caption = "Supplementary Material: Table S1 - Spearman's correlations for monolingual and multilingual children in the Polish dataset", placement = "H")

```

```{r mse}
# TODO
# 1. Eng values shifted slightly? 

# Polish
by_child_pl <- by_child_pl %>% mutate(sq_err = (full_theta - cat_theta)^2,
                                      full_cat_diff = full_theta - cat_theta)
mean(by_child_pl$sq_err) # 0.19
median(by_child_pl$sq_err) # 0.08
sd(by_child_pl$sq_err) # 0.45

bad_thresh_pl <- mean(by_child_pl$sq_err) + 1.5*sd(by_child_pl$sq_err)
bad_thresh_pl # 0.86

by_child_pl <- by_child_pl %>% mutate(extreme_discrep = 
                                           case_when(
                                                sq_err > bad_thresh_pl ~ "yes",
                                                .default = "no"
                                           ))

table(by_child_pl$extreme_discrep)

# English

# mean squared error
prod_s <- prod_s %>% mutate(sq_err = (fullTheta - catTheta)^2,
                            full_cat_diff = fullTheta - catTheta)

bad_thresh_en <- mean(prod_s$sq_err) + 1.5*sd(prod_s$sq_err) # mean=.55, med = .17

# bad_en <- subset(prod_s, sq_err > bad_thresh_en) # 18 --> 15?
# good_en <- subset(prod_s, sq_err <= bad_thresh_en) # 186 --> 189?

prod_s <- prod_s %>% mutate(extreme_discrep = 
                                           case_when(
                                                sq_err > bad_thresh_en ~ "yes",
                                                .default = "no"
                                           ))

table(prod_s$extreme_discrep)


```

```{r duration-full}

# Polish
mean(by_child_pl$duration_full) # 133892.9 sec
sd(by_child_pl$duration_full) # 980822.7 sec
# sd larger than mean - highly skewed

ecdf_fun <- ecdf(as.numeric(by_child_pl$duration_full))
by_child_pl <- by_child_pl %>%
  mutate(duration_full_percentile = ecdf_fun(duration_full) * 100)

# English
# create and attach duration_full to prod_s

demo <- demo %>% mutate(
     duration_full = lubridate::as.duration(last_modified_full - created_date_full))
prod_s <- merge(prod_s, demo[,c("subject_id", "duration_full")], all.x = T, by = "subject_id")

ecdf_fun <- ecdf(as.numeric(prod_s$duration_full))
prod_s <- prod_s %>%
  mutate(duration_full_percentile = ecdf_fun(duration_full) * 100)

prod_s %>% ggplot(aes(y = duration_full, x = extreme_discrep)) + geom_jitter(alpha = 0.3)
# extreme_discrep cases are scattered across the whole range of durations

```

We also looked at the mean squared error between the abilities as estimated by CAT-CDI and from the full CDI. The mean squared error in English was 0.55 (*Mdn* = 0.17, *SD* = 1), and in Polish it was `r mean(by_child_pl$sq_err)` (*Mdn* = `r median(by_child_pl$sq_err)`, *SD* = `r sd(by_child_pl$sq_err)`). We also looked at the children for whom the estimates from the CAT-CDI and full CDI diverged extremely, i.e. their difference between the errors was 1.5 SD from the mean. There were `r sum(prod_s$extreme_discrep == "yes")` such cases (`r round(sum(prod_s$extreme_discrep == "yes")/count(prod_s)*100,2)`%) in the English dataset and `r sum(by_child_pl$extreme_discrep == "yes")` cases (`r round(sum(by_child_pl$extreme_discrep == "yes")/count(prod_s)*100,2)`%) in the Polish dataset. All participants in both datasets showed higher ability estimates on the CDI-CAT compared to the full CDI. If the full CDI is considered the baseline, this suggests that parents may have overestimated their child’s vocabulary on the CDI-CAT, potentially responding “yes – produces” to more items than expected based on full CDI estimates (as suggested by Kachergis, et al. 2022). An alternative explanation is that, for these participants, the full CDI may have underestimated the child’s true ability. Notably, all Polish participants with large discrepancies completed the full CDI in unusually short times (their completion times were among the shortest 5% in the sample) suggesting their responses may have been rushed or less attentive. This could have led to lower ability estimates from the full CDI. Supporting this interpretation, their CDI-CAT scores had acceptable measurement errors (below or equal to 0.1 for Polish), indicating reliable ability estimation by the CDI-CAT, in contrast to the full CDI. However, this pattern did not appear in the English dataset, where only `r sum(by_child_pl$extreme_discrep == "yes" & by_child_pl$duration_full_percentile <= 5)` participants who showed extreme discrepancy also showed very short administrations of the full CDI. 

```{r, include = T}
# participants with extreme discrepancy in English dataset:
extreme_discrep <- prod_s %>% filter(extreme_discrep == "yes") %>% select(-subject_id, duration_full_percentile)
apa_table(extreme_discrep, landscape = T)
```

```{r}
prod_s %>% filter(extreme_discrep == "no") %>% summarise(mean(sq_err)) # 0.44, was 0.55
by_child_pl %>% filter(extreme_discrep == "no") %>% summarise(mean(sq_err)) # 0.12, was 0.19
```

We also re-calculated the mean squared error without the cases of extreme discrepancy, which yielded a MSE of `r round(prod_s %>% filter(extreme_discrep == "no") %>% summarise(mean(sq_err)),2)` (*Mdn* = `r round(prod_s %>% filter(extreme_discrep == "no") %>% summarise(median(sq_err)),2)`, *SD* = `r round(prod_s %>% filter(extreme_discrep == "no") %>% summarise(sd(sq_err)),2)`) in English and MSE of `r round(by_child_pl %>% filter(extreme_discrep == "no") %>% summarise(mean(sq_err)),2)` (*Mdn* = `r round(by_child_pl %>% filter(extreme_discrep == "no") %>% summarise(median(sq_err)),2)`, *SD* = `r round(by_child_pl %>% filter(extreme_discrep == "no") %>% summarise(sd(sq_err)),2)`) in Polish.

## Item properties in the two CAT-CDIs
Our second aim was to analyze similarities and differences in IRT item properties and item selection in CAT in English and Polish.

There are 679 items in the English CAT-CDI and 666 items in the Polish CAT-CDI. For both sets of items, the items' difficulty and discrimination parameters were calculated using IRT 2 parameter model (these included separate samples, see Kachergis et al. 2022 and Krajewski et al. (in preparation)). An item’s difficulty indicates the ability level at which there is a 50% probability that a participant will respond correctly <!-- SEE ISSUE 1: https://github.com/karolina-m/catcdi_validation_eng_pl/issues/1 -->. It is on the same scale as ability with negative values indicating difficulty items, values around 0 indicating medium difficulty, and positive values indicating easy items. <!-- we should probably add a footnote here explaining that d is kind of inversely related to theta and so negative ability means low while positive ability means high. --> An item’s discrimination indicates how well it distinguishes between individuals with slightly different ability levels--especially those near that difficulty point. Of these two parameters, item difficulty is of greater interest to the present paper as it is directly linked to ability and as discrimination power is more about how good the item is at measuring, rather than what it is measuring.<!-- this could potentially go up, into Intro/Methods? -->

```{r diff_across_lgs}
# item difficulty by language 
items_cat_en %>% 
     dplyr::summarise(min(d), max(d), mean(d), median(d), sd(d))

items_cat_pl %>% 
     dplyr::summarise(min(d), max(d), mean(d), median(d), sd(d))

t1 <- apa_print(t.test(items_cat_en$d, items_cat_pl$d))

```

```{r common_items}
# common items - still different mean difficulty?

items_en_pl <- merge(items_cat_pl, items_cat_en, 
                  all.x = T, all.y = T,
                  suffixes = c("_pl","_en"),
                  by = "uni_lemma") %>% filter(!is.na(item) & !is.na(definition))
count(items_en_pl) # 390

t2 <- apa_print(t.test(items_en_pl$d_en, items_en_pl$d_pl, paired = T))
# t = -23.112, df = 389, p-value < 2.2e-16

# visualise
items_en_pl %>% select(uni_lemma, d_pl, d_en) %>% pivot_longer(cols = d_pl:d_en, names_to = "lang", values_to = "d") %>% ggplot(aes(x = lang, y = d, group = lang)) + geom_boxplot() + geom_jitter(alpha = 0.3)

# corr
c <- apa_print(cor.test(items_en_pl$d_pl, items_en_pl$d_en)) # 0.65
```

English items are more difficult than the Polish items, `r t1$full_result` (English: min = `r round(min(items_cat_en$d),2)`, max = `r round(max(items_cat_en$d),2)`, *M* = `r round(mean(items_cat_en$d),2)`, *Mdn* = `r round(median(items_cat_en$d),2)`, *SD* = `r round(sd(items_cat_en$d),2)`; Polish: min = `r round(min(items_cat_pl$d),2)`, max = `r round(max(items_cat_pl$d),2)`, *M* = `r round(mean(items_cat_pl$d),2)`, *Mdn* = `r round(median(items_cat_pl$d),2)`, *SD* = `r round(sd(items_cat_pl$d),2)`). Notably, this was true even for a subset of `r count(items_en_pl)` items common to both languages - these items still proved to be more difficult in English than in Polish: `r t2$full_result`. This difference in mean difficulty may be influenced by the characteristics of the samples used to estimate the IRT models. In English, item difficulty was calculated based on a broader sample of children aged 12–36 months (spanning the CDI:WG, CDI:WS, and CDI-III), whereas the Polish data came from a sample of narrower age range of 18–36 months, corresponding to the CDI:WS. As a result, item difficulty in English was estimated using a relatively younger sample, for whom certain items may have been more challenging—thus appearing more difficult—compared to the older Polish sample. Still the item difficulty for common items in the two languages was positively and moderately correlated: `r c$full_result`.

```{r d_by_cdi_category_tab, include = T}
# correlations by cdi semantic category on common words
# spearman because ranked

items_en_pl %>% group_by(category_pl) %>% summarise(
     rho = cor(d_pl, d_en, method = "spearman", use = "complete.obs"),
     n = length(category_pl)) %>% arrange(rho) -> corr_by_category

apa_table(corr_by_category, placement = "H")
```

```{r d-by-cdi-category-fig, fig.width = 10, fig.height = 5, fig.cap = "The relative positioning of the items by semantic category (colored) plotted in the context of the full item pool (grey): (A) English, (B) Polish.", fig.pos = '!h', include = T}

# TODO: adjust order, put "helping_verbs" at the end

# AMERICAN ENGLISH
fig1a <- items_cat_en %>%
ggplot(aes(d, a1))  + 
  geom_point(data = items_cat_en %>%
               select(-category), aes(x = d, y = a1), color = 'grey', size = 0.4) +
  geom_point(aes(color=as.factor(category)), size = 0.4) +
  facet_wrap(~category) +
  xlim(-8, 8) +
  ylim(0,8) +
  labs(x = "Item difficulty (d)", y = "Item discrimination (a1)", title = "(A) English") +
  theme(legend.position = "none")

# TODO: adjust order, put "modal_adverbs" and "prepositions" at the end

# POLISH
fig1b <- items_cat_pl %>%
  mutate(category = case_when(
    category == "descriptive_words (adjectives)" ~ "descriptive_words",
    category == "descriptive_words (adverbs)" ~ "descriptive_words",
    category == "pronouns_demonstrative" ~ "pronouns",
    category == "pronouns_personal" ~ "pronouns",
    .default = category
  )) %>%
  filter(!is.na(category)) %>% # tylko has category == NA
ggplot(aes(d, a1))  + 
  geom_point(data = items_cat_pl %>%
               select(-category), aes(x = d, y = a1), color = 'grey', size = 0.4) +
  geom_point(aes(color=as.factor(category)), size = 0.4) +
  facet_wrap(~category)  +
  xlim(-8, 8) +
  ylim(0,8) +
  labs(x = "Item difficulty (d)", y = "Item discrimination (a1)", title = "(B) Polish") +
  theme(legend.position = "none")

fig1 <- ggarrange(fig1a, fig1b, ncol = 2, nrow = 1)
fig1

```

We also wanted to do check whether items in particular CDI semantic categories show related parameter values across the two languages. We performed a series of Spearman's rank correlations (on the items common to CDI-CATs in Polish and English) for each CDI semantic category (see Table \@ref(tab:d_by_cdi_category_tab)). The rank correlations coefficients vary by category, but for half of the categories the correlations are moderate to strong (0.52 to 0.91). Figure \@ref(fig:d-by-cdi-category-fig) shows the relative positioning of the items by semantic category (colored) plotted in the context of the full item pool (grey) in the two languages. It can be seen from the figure that many categories (e.g., sounds) show a similar distribution of items relative to the whole item pool across Polish and English.

## Item selection in the two CAT-CDIs

```{r}
# Data prep

# POLISH
items_run_pl <- responses_cat_pl %>% select(items) %>% group_by(items) %>% summarise(n = length(items))
items_run_pl <- merge(items_cat_pl, items_run_pl, by.x = "item", by.y = "items", all.x = T)
items_run_pl[is.na(items_run_pl)] <- 0
items_run_pl$never <- ifelse(items_run_pl$n == 0, "never", "at_least_once")

items_run_pl <- items_run_pl %>% 
  mutate(items_perc = (n/115)*100) %>%
  mutate(percent_for_color = ifelse(items_perc == 0, NA, items_perc)) 

# AMERICAN ENGLISH

items_run_en <- resps %>%
  select(subject_id, response_cat, definition)

items_run_en <- items_run_en %>% mutate(
  item_used = case_when(
    response_cat == "no_test" ~ 0,
    .default = 1
  )
)

items_run_en <- items_run_en %>%
  select(definition, item_used) %>%
  group_by(definition) %>%
  mutate(n = sum(item_used)) %>%
  select(definition, n) %>% unique()

# how many participants in resps?
resps %>% select(subject_id) %>% unique() %>% nrow()

items_run_en <- items_run_en %>% 
  mutate(items_perc = (n/204)*100) %>%
  mutate(percent_for_color = ifelse(items_perc == 0, NA, items_perc)) 

items_run_en <- merge(items_run_en, coefs_2pl[,c("definition", "a1", "d")], 
                   all.x = T,
                   by = "definition")

# ame items lack categories
eng_ws_items <- read_csv("data/english/eng_ws_items.csv")
items_run_en <- merge(items_run_en, eng_ws_items[,c("definition", "category")], all.x = T, by = "definition")
rm(eng_ws_items)
```

```{r}
# PL
sum(items_run_pl$n != 0) # 258
round(sum(items_run_pl$n != 0)/count(items_run_pl)*100,2) # 38.74

# EN
sum(items_run_en$n != 0) # 251
round(sum(items_run_en$n != 0)/count(items_run_en)*100,2) # 36.91
```
```{r used-items-fig, fig.width = 10, fig.height = 5, fig.cap = "How often a given item was used in CAT-CDI administrations in the validation study in (A) English and (B) Polish. The items (points) are colored by the percentage of their appearance in the CAT-CDI administrations. Items colored in grey are items never used in any of the CAT-CDI administrations.", fig.pos = '!h', include = T}

# Plotting (un)used items - cloud
# PL
fig2b <- items_run_pl %>%
  mutate(d_bin = cut(d, breaks = seq(-7.5, 7.5, by = 1))) %>%
  ggplot(aes(x = d_bin, y = a1)) +
  geom_jitter(aes(color = percent_for_color), alpha = 0.7, width = 0.2) +  # Points
  geom_boxplot(alpha = 0.2, width = 0.5, outlier.shape = NA) +  # Add boxplots
  scale_color_viridis(option = "D", na.value = "grey", name = "Percent", begin = 1, end = 0) +
  ylim(0, 7.5) +
  labs(
    x = "Difficulty (d)",
    y = "Discrimination (a1)",
    title = "(B) Polish",
    caption = "Grey are unused items"
  ) +
  theme_light() +
  theme(legend.position = c(0.95, 0.85))

# EN
fig2a <- items_run_en %>%
  mutate(d_bin = cut(d, breaks = seq(-7.5, 7.5, by = 1))) %>%
  ggplot(aes(x = d_bin, y = a1)) +
  geom_jitter(aes(color = percent_for_color), alpha = 0.7, width = 0.2) +  # Points
  geom_boxplot(alpha = 0.2, width = 0.5, outlier.shape = NA) +  # Add boxplots
  scale_color_viridis(option = "D", na.value = "grey", name = "Percent", begin = 1, end = 0) +
  ylim(0, 7.5) +
  labs(
    x = "Difficulty (d)",
    y = "Discrimination (a1)",
    title = "(A) English",
    caption = "Grey are unused items"
  ) +
  theme_light() +
  theme(legend.position = c(0.95, 0.85))
fig2 <- ggarrange(fig2a, fig2b, nrow = 1, ncol = 2)
fig2
  

```


It is to be expected that a CDI-CAT will not need to administer all the items in the item bank. In fact, in the validation study the English CAT-CDI used `r sum(items_run_en$n != 0)` items (`r round(sum(items_run_en$n != 0)/count(items_run_en)*100,2)`%) and similarly, the Polish CAT-CDI used `r sum(items_run_pl$n != 0)` items (`r round(sum(items_run_pl$n != 0)/count(items_run_pl)*100,2)`%).
By design, a CDI-CAT selects items that are most informative for each participant. This means it draws from a subset of available items--typically those matched to the participant’s current ability estimate in terms of difficulty, and with high discrimination, meaning they effectively distinguish between individuals with abilities close to that estimate. Figure \@ref(fig:used-items-fig) plots the items by their difficulty and discrimination parameters and colors the items (points) by how often they were used in CAT-CDI administrations in English and Polish. A few findings are of note here. First, both CAT-CDIs used items of very low discrimination (i.e, items that do no discriminate well between ability levels). Often that was because for a given difficulty level there were not enough items and CAT-CDI had to use all the items available (this is particularly true of the items on the two ends of difficulty). However, more surprisingly, it is also true of items of medium difficulty, where items of higher discrimination were available (but were not chosen by the CAT algorithm).
<!-- SEE ISSUE #2 https://github.com/karolina-m/catcdi_validation_eng_pl/issues/2. NOTE TO US: I emailed Phil Chalmers, the author of mirtcat and mirt packages in R to ask if he has any idea why this is so -->

Finally, some items keep being shown in many CAT administrations. In English, 3 items appear in more than half of administrations - *"long"* in 64% of administrations, *"make"* in 54% of administrations, and *"last"* in 53% of administrations. They are also high discrimination items in their difficulty bins. "Long" is additionally one of the starting items, i.e. a first item shown to the parent, chosen so that there is high probability the parent can respond positively. In Polish, 3 items appear in more than half of administrations - *"szukać"* (to look for) appearing in 83% of administrations, *"znaleźć"* (to find) appearing in 61% of administrations, and *"babcia"* (grandmother) appearing in 60% of administrations. All three items show very high discrimination. "Szukać" (to look for) and "znaleźć" (to find) are also in top 5 discrimination items in general. "Babcia" (grandmother) is one of the starting items. 

```{r}
items_en_pl <- merge(items_en_pl, items_run_pl[,c("item", "items_perc")], all.x = T, by = "item")
names(items_en_pl)[names(items_en_pl) == "items_perc"] <- "items_perc_pl"

items_en_pl <- merge(items_en_pl, items_run_en[,c("definition", "items_perc")], all.x = T, by = "definition")
names(items_en_pl)[names(items_en_pl) == "items_perc"] <- "items_perc_en"

c3 <- apa_print(cor.test(items_en_pl$items_perc_pl, items_en_pl$items_perc_en, paired = T, method = "spearman")) # .37

```


To check whether the CDI items common in both languages may show similar frequencies in CAT-CDI administrations, we ran Spearman's rank correlation and found that the items frequencies in administrations were weakly correlated, `r c3$full_result`. However, the two CAT-CDIs differed in their potential length, as the English CAT-CDI was set to administer 25--50 items and the Polish CAT-CDI could administer up to 75 items (no minimum was set, but in practice the minimum items administered were 13). As an item is more likely to appear in longer tests (simply due to test length), then frequency of the Polish items could be inflated. <!-- Could it be that, as an item is more likely to appear in longer tests (simply due to test length), frequency of the Polish items could be inflated? Could it influence the item's rank? I could probably recalculate item frequency not per total number of administrations, but per total number of items administered? SEE ISSUE #3 https://github.com/karolina-m/catcdi_validation_eng_pl/issues/3  -->

```{r}

```


## CAT-CDIs' usefullness in research and practice



# Discussion

1. CAT-CDIs are strongly correlated with the full CDI in both languages. That's a good sign for the overall CAT development. Also the correlations are high for both monolingual and multilingual sample, though it's a small sample in Polish and more research is needed (and planned).    
2. CAT-CDIs showed diverged largely from full estimates only for a handful of children. We cannot yet rule out that CAT was to blame, but for large majority of them (with few exceptions) the ability estimate was reliable (acceptable SE). Could it be 
4. Item properties - correlated (moderately!) across lgs, differences due to calibration samples? similarities across lgs per cdi semantic category
5. 

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::

`r cite_r("r-references.bib")`
