---
title             : "Short Adaptive CDIs: Validation and Cross-Linguistic Comparison of Computer-Adaptive Parent Report Measures of Early Vocabulary for Polish and English"
shorttitle        : "Short Adaptive CDIs: Validation and Cross-Linguistic Comparison"

author: 
  - name          : "Karolina Muszyńska"
    affiliation   : "1"
    corresponding : yes    
    address       : "University of Warsaw, Faculty of Psychology, ul. Stefana Banacha 2D, 02-097 Warsaw, Poland"
    email         : "karolina.muszynska@psych.uw.edu.pl"
    role: # Contributorship roles from here: https://docs.google.com/spreadsheets/d/1Gxq875LTPtNXkmpYBppRpVkog8T00eZfETX5Lp1kOck/edit?gid=0#gid=0 
      - "Conceptualization"
      - "Data Curation"
      - "Formal Analysis"
      - "Funding Acquisiton"
      - "Investigation"
      - "Methodology"
      - "Project administration"
      - "Software"
      - "Supervision"
      - "Validation"
      - "Visualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Grzegorz Krajewski"
    affiliation   : "1"
    role:
      - "Conceptualization"
      - "Data Curation"
      - "Investigation"
      - "Methodology"
      - "Project administration"
      - "Software"
      - "Supervision"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Magdalena Krysztofiak"
    affiliation   : "1"
    role:
      - "Conceptualization"
      - "Data Curation"
      - "Funding Acquisiton"
      - "Investigation"
      - "Methodology"
      - "Supervision"
      - "Visualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "George Kachergis"
    affiliation   : "2"
    role:
      - "Conceptualization"
      - "Data Curation"
      - "Investigation"
      - "Methodology"
      - "Software"
      - "Supervision"
      - "Writing - Review & Editing"
  - name          : "Virginia A. Marchman"
    affiliation   : "3"
    role:
      - "Conceptualization"
      - "Supervision"
      - "Writing - Review & Editing"
      - "Supervision"
  - name          : "Michael C. Frank"
    affiliation   : "3"
    role:
      - "Conceptualization"
      - "Supervision"
      - "Project administration"
      - "Writing - Review & Editing"

affiliation:
  - id            : "1"
    institution   : "Faculty of Psychology, University of Warsaw, Poland"
  - id            : "2"
    institution   : "Cognition, Behavior, and Information (CBI), Arizona State University, USA"
  - id            : "3"
    institution   : "Department of Psychology, Stanford University, USA"

abstract: |
  Abstract here
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : "References from.bib"

floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}

library(papaja)
r_refs("References from.bib") #exported as bibtex from https://docs.google.com/document/d/1MPInUTyY2Y81EtiiK0tUOo8vG3zowfqiXjpBF0p8Mzs/edit?tab=t.0

library(tidyverse)
library(Multilada)
library(here)
library(effectsize)


```

```{r analysis-preferences}
# Seed for random number generation
set.seed(123)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = F, include = F, warning = F, message = F, out.width="100%")
```


```{r, echo = F}
# Data
source(here("data/polish/data_polish.R"))
source(here("data/english/data_ame_english.R"))
```
# Introduction

MacArthur-Bates Communicative Development Inventories (CDI) are parent-report forms that are used worldwide in research and clinical practice to measure children’s language abilities, especially word knowledge. There are over 100 language adaptations of the CDI, including Polish (Smoczyńska et al. 2015), and up to three age versions in each language: Words and Gestures (CDI:WG) for children up to the age of a year and a half, Words and Sentences (CDI:WS) for older children up to around three, and CDI-III for children from three to three and a half or four (exact age ranges vary between adaptations). All CDIs include a vocabulary checklist, a list of words in which parents or other caregivers mark the words that their child “understands” and/or “understands and says.” CDIs and similar parental reports draw on caregivers’ extensive experiences with the child in a variety of contexts. They demonstrate adequate validity with spontaneous language samples (Gatt et al. 2023; O' Toole and Fletcher 2010) and direct language measures (e.g. Dale 1991; Lemhöfer and Broersma 2011; Smolík and Bytešníková 2021) both in typically-developing children as well as children with an elevated likelihood or diagnosis of ASD (Belteki et al. 2025). 

The vocabulary checklists include many words across numerous semantic and lexical categories, varying in their frequency, age-appropriateness, and other properties (e.g., Frank et al. ; Garmann et al. 2019). As a result, CDIs tend to include hundreds of words and require a substantial amount of time (e.g., 20-30 minutes) for caregivers to complete. The initial attempts to shorten the CDI word lists were to simply identify a subset of items from the original lists that correlated most strongly with responses on the full list (Fenson et al. 2000; Ezeizabarrena et al. 2024; Smolík and Bytešníková 2021; Lasorsa et al. 2021; Kim et al. 2014). More recently, Mayor and Mani (2019) administered a random subset of words from the full CDI, combining these with vocabulary data from the WordBank database, and then incorporated information about the child’s age, sex, and language status into Bayesian models that predict the child’s overall vocabulary size as it would appear on the full CDI. This method provided a reliable estimate of the child’s language development level without requiring the complete inventory: using a 25-item lists, CDIs correlated with the full CDIs at r = .96, with an SE of .14 and a reliability of .98; using a 50-items CDIs showed slightly lower correlation at r = .94 with similar SE of .14 and a reliability of .98. However, as noted by Kachergis et al. (2022), while using age and sex to predict vocabulary improves precision with limited data, these factors can also bias individual assessments by pulling scores toward typical expectations for those demographics.

Other attempts to shorten the CDIs have incorporated Item Response Theory (IRT) and Computerized Adaptive Testing (CAT) (e.g., Makransky et al. 2016; Chai et al. 2020; Kachergis et al. 2022). IRT is a statistical framework used in educational and psychological testing to model the relation between individuals’ responses to test items and their underlying abilities. Unlike Classical Test Theory (CTT), which focuses on the overall test score (and relates that to an individual’s ability), IRT examines each response and how it reflects the tested ability. The measured ability (often denoted as theta) is placed on a continuum: negative values correspond to lower ability, values near zero to average ability, and positive values to higher ability. The focus of the IRT is on the items’ characteristics (or parameters). In a 2-parameter IRT model (applied here), each item is characterised in terms of its difficulty (the ability level required to know the item) and discrimination (how effectively the item distinguishes between similar ability levels). If we know the items’ characteristics, every response to an item can provide some information about the individual’s theta. In the CDI context, IRT models the CDI items’ characteristics (i.e. difficulty and discrimination), and the relation between parental responses to individual items and the child’s measure ability, i.e. vocabulary knowledge. CAT is a form of computer-based testing that builds on IRT. In a CAT-CDI, the CDI items are shown to the parent one at a time. With each response to an item of given parameters, the algorithm updates the estimate of the child’s ability (i.e., vocabulary knowledge) and selects the next CDI item to refine this estimate. The process continues until a set number of items has been shown or the estimate reaches an acceptable level of precision. IRT and CAT-based tasks are being developed in many areas of language assessment, e.g. reading (Ma et al. 2025), direct vocabulary skills in older children (Bohn et al. 2024; Bohn et al. 2025) and measuring overall variability in children's learning and development (Frank et al. 2025).

The first IRT-based CDI was developed by Makransky et al. (2016) who applied IRT models to normative data from the CDI:WS to create computer adaptive test (CAT) versions. Using the full CDI dataset, they simulated CAT-CDIs of different lengths by sequentially selecting items and filling in responses based on participants’ complete CDI data. Then they compared the full CDI scores with the scores from various fixed-length CATs ranging from 5 to 400 items. Notably, a CAT with just 50 items produced scores that correlated strongly with the full CDI, achieving a correlation of r = .95. Building on this, Kachergis et al. (2022) developed online CATs for assessing vocabulary in both English and Mexican Spanish and extended the adaptive assessment framework to both receptive and productive vocabulary. Their CAT-CDIs were found to accurately estimate vocabulary comprehension and production using as few as 25–50 items, significantly reducing assessment time while maintaining high validity. These developments laid the groundwork for the creation of a CAT-CDI in Polish (Krajewski et al., in preparation). 

Both in CAT-CDIs in English (Kachergis et al. 2022) and Polish (Krajewski et al., in preparation) we used a two-parameter logistic (2PL) model, in which each CDI item is defined by a difficulty parameter (reflecting the ability level required to know the item) and a discrimination parameter (indicating how well the item distinguishes between individuals with similar ability levels). These two item characteristics were modelled on large datasets: Wordbank data in English, and CDI:WS norming data in Polish.

Once an IRT model is established and item parameters are estimated, a computerized adaptive task (CAT) can be constructed. This involves specifying how the task should begin, how the ability is to be estimated, and the criteria for ending the task. In both the English and Polish CAT-CDIs, basic caregiver-reported information – specifically the child’s age and sex – is used to generate an initial ability estimate. The first item is then selected based on this estimate. In both versions, we deliberately present a slightly easier item (below the estimated ability level) to encourage a positive response to the first item. Parental response to one item at a time allows the CAT algorithm to update the ability estimate and calculate a standard error (SE) of that estimate. The task can end when a maximum number of items (e.g. 50) is shown to the caregiver, or when a particularly low error of the estimate is reached.

Importantly, when developing a CAT-CDI in a new language, researchers must make several crucial decisions. In what follows, we illustrate these decisions using the English and Polish CAT-CDIs as examples. 

One key decision is how many CAT-CDIs should be developed. For English, the CAT-CDIs are divided into two instruments: one for production and one for comprehension. Unlike the original CDIs, there are no separate age-based versions; instead, both English CAT-CDIs cover the full developmental span of the CDI, from the earliest CDI:WG through the CDI-III. This approach assumes that early vocabulary acquisition can be represented by two latent traits – comprehension and production – and maximizes the dataset used to model them. As a result, only two English CAT-CDIs exist, which is also advantageous for practitioners: they do not need to determine which version (WG, WS, or CDI-III) to administer for a given child’s age. In contrast, the Polish CAT-CDIs preserve the structure of the original instruments. Separate CAT-CDIs were developed for each version, including the CAT-CDI:WS (word production). 

This structural choice has important consequences for how item characteristics are estimated. The English CDI item parameters – including item difficulty – were estimated using a sample of children aged 12 to 36 months (see Kachergis et al., 2022, for details), covering the entire age range of the CDI instruments. By contrast, the Polish CDI items in the CAT-CDI:WS were calibrated on a narrower range of 18 to 36 months, consistent with the norming age range for the Polish CDI:WS. As a result, an item may appear more difficult in the English CAT-CDI (compared to Polish) simply because its parameters were estimated using responses from a younger and more heterogeneous sample. At the same time, we can still expect cross-linguistic similarities – both at the item level (e.g., the word “mommy” is likely to be easy in both Polish and English) and at the category level (e.g., sound-related words tend to be generally easier across languages).

Another important decision concerns the stopping criteria of the CAT task. These criteria can be based on a maximum number of administered items or reaching a target level of estimate precision – typically expressed as a standard error (SE) of the estimate. The English CAT-CDI prioritized a shorter CAT typically ending after 50 items, or sooner if a standard error (SE) of 0.15 was reached. In contrast, the Polish CAT-CDI prioritised greater precision, continuing to a maximum of 75 items until an SE of 0.1 was reached. If this level of precision could not be reached, the test concluded after a maximum of 75 items. These differences in CAT settings may result in relatively shorter English CAT-CDIs but relatively more precise Polish CAT-CDIs.


The aim of this paper is to evaluate the potential usefulness of CAT-CDIs for both research and clinical applications, using evidence from the validation studies of the two CAT-CDIs in English and in Polish. For a CAT-CDI to be effective in these contexts – whether for screening, diagnosis, or broader research purposes – it should meet the following criteria:
1. Require less administration time.     
2. Produce estimates that align closely with scores obtained from the full CDI.      
3. Maintain high measurement precision (i.e., low standard error) across the relevant ability range.      
4. Avoid repeated reliance on the same subset of items, which is particularly important in longitudinal or repeated-measures designs.
We will report these outcomes in the two CAT-CDIs, in English and in Polish and discuss similarities and differences in the context of how the development of two CAT-CDIs differed. We will also discuss how CAT-CDI results (an ability estimate) can be translated to a clinical context. 

We will report these outcomes in the two CAT-CDIs, in English and in Polish and discuss similarities and differences in the context of how the development of two CAT-CDIs differed. The English data presented here were also included in a brief validation report by Kachergis et al. (2022). In the present paper, however, we examine these metrics in greater detail and place them in relation to the validation results from the Polish CAT-CDI.

# Methods

## Participants

```{r}
# the two groups differ by age
t_age <- apa_print(t.test(by_child_pl$age_CAT, kept_demo$age_cat))
```


In the American English validation study, the participants were 204 parents of children aged 15 to 36 months (*M* = `r round(mean(kept_demo$age_cat),2)`, *Mdn* = `r round(median(kept_demo$age_cat),2)`, SD = `r round(sd(kept_demo$age_cat),2)`). There were 103 girls in the sample (`r round(sum(kept_demo$sex_cat == "F")/count(kept_demo)*100,2)`%). The sample was ethnically diverse, including White, Black, Latino/Hispanic, Asian, Native American, and other races/ethnicities. None of the children were reported to hear other languages apart from English. <!-- as checked in kept_demo$other_languages --> The average
number of years of education attained by the primary caregiver was `r round(mean(kept_demo$primary_education_cat),2)` (SD = `r round(sd(kept_demo$primary_education_cat),2)`).

In the Polish validation study, the participants were 113 parents of children aged 18 to 34 months (*M* = `r round(mean(by_child_pl$age_CAT),2)`, *Mdn* = `r round(median(by_child_pl$age_CAT),2)`, SD = `r round(sd(by_child_pl$age_CAT),2)`). There were 49 girls in the sample (`r round(sum(by_child_pl$sex == "dziewczynka")/count(by_child_pl)*100,2)`%). The sample was White (reflecting the limited ethnic diversity within the Polish population). However, there were 28 children multilingual with Polish (as their home language) and some other (societal) language, living outside of Poland. Parents were asked about their education level and `r sum(by_child_pl$parental_edu == "HIGHER_EDU" | by_child_pl$parental_edu == "PHD", na.rm = T)` reported having higher education or PhD, `r sum(by_child_pl$parental_edu == "UNFINISHED_HIGHER_EDU", na.rm = T)` reported having unfinished higher education, `r sum(by_child_pl$parental_edu == "SECONDARY", na.rm = T)` reported having secondary education and `r sum(is.na(by_child_pl$parental_edu))` did not report their education level. 

Notably, the two validation samples differed slightly (but significantly) in the mean age of children, `r t_age$full_result`, which may affect the outcomes of direct comparisons between the samples, where age could be a contributing factor. We present one such comparison--standard error of CAT ability--where this difference is explicitly discussed. *NOTE*: You can also view/comment on this on GitHub [Issue #7 linked here](https://github.com/karolina-m/catcdi_validation_eng_pl/issues/7).


## Materials
The data was gathered with CAT-CDIs in American English and Polish. The American English CAT-CDI was created from the items that were common to three CDIs: Words and Gestures, Words and Sentences and CDI-III. With these items, two CAT-CDIs were created, one for word production and another for word comprehension (see Kachergis et al. 2022 for more details). The Polish CAT-CDIs were created following a largely similar procedure as the ones in American English, but there was a separate CAT-CDI developed for each CDI version (CDI: Words and Gestures and CDI: Words and Sentences) and for each CDI:WG subscale (word production, word comprehension, gesture use) (Krajewski et al. in preparation). The American English CAT-CDI word production included 679 items and the Polish CAT-CDI version of WS: Words and Sentences included 666 items. There are 395 words common to both CAT-CDI language versions. 

## Procedure
Though in both American and Polish validation studies the parents were asked to fill in both the full CDI and the CAT-CDI, the procedure differed between the two studies. In the validation of American English CAT-CDI, parents were given both CDIs in one sitting, but the order of the CDI versions was counterbalanced. In the Polish validation study, the order of the CDIs was fixed, with the full CDI always first, and then, after 1–30 days (*M* = `r round(mean(by_child_pl$days_between),2)`, *Mdn* = `r round(median(by_child_pl$days_between),2)`, SD = `r round(sd(by_child_pl$days_between),2)`) parents were asked to fill in the CAT-CDI. Both studies were done fully online and obtained the approval of adequate ethics committees.

# Results

We assume that, for a CAT-CDI to be useful in the contexts of research and practice (e.g., for screening or diagnosis), it should meet the following criteria:   
1.	Take less time.    
2.   Its estimates should relate closely to the scores and estimates obtained from the full CDI.   
3.	Maintain high precision (i.e., low standard error) across relevant levels of ability.   
5.	Avoid overuse of the same subset of items—particularly important in repeated measures with the same participants (e.g., in longitudinal studies).      
Additionally, we would like to compare the item difficulty in the two CAT-CDIs

In the following section we will explore the two CAT-CDI's performance on each of those criteria.


## Length of administration
```{r}
# calculate and add duration_cat for Eng
demo <- demo %>% mutate(
     duration_cat = lubridate::as.duration(last_modified_cat - created_date_cat))

prod_s <- merge(prod_s, demo[,c("subject_id", "duration_cat")], all.x = T, by = "subject_id")

```

```{r}
# mean number of items administered in CAT
# PL
by_child_pl %>% summarise(mean(no_items), #35
                          median(no_items),#23
                          min(no_items), #13
                          max(no_items), #75
                          sd(no_items)) #22.72

# EN
resps_sum <- resps %>% filter(response_cat != "no_test") %>%
     group_by(subject_id) %>%
     summarise(n = length(test_word))

resps_sum %>% summarise(mean(n), #31
                          median(n), #25
                          min(n), #25
                          max(n), #50
                          sd(n)) #10.11

t1 <- apa_print(t.test(resps_sum$n, by_child_pl$no_items))
```

One of the key metrics for CAT-CDI's usefulness is whether it can shorten the CDI administration compared to the full CDI (but still obtain a reliable estimate of child's lexical ability). We have found that CAT-CDIs were significantly shorter to administer: the English CAT-CDI median time was `r duration(median(prod_s$duration_cat))` (median full CDI time was `r duration(median(prod_s$duration_full))`) and the Polish CAT-CDI median time was `r duration(median(by_child_pl$duration_cat))` (median full CDI time was `r duration(median(by_child_pl$duration_full))`). Even though the CAT-CDI in Polish seemed shorter (compared to English CAT-CDI), the number of items administered in the two language versions of CAT-CDI was comparable: English *M* = `r round(mean(resps_sum$n),2)`, *Mdn* = `r median(resps_sum$n)` (25--50 items), Polish *M* = `r round(mean(by_child_pl$no_items),2)`, *Mdn* = `r median(by_child_pl$no_items)` (`r min(by_child_pl$no_items)`--`r max(by_child_pl$no_items)` items).

## Comparison of estimates

### Correlations

Our first aim was to examine whether CAT-CDIs in American English and Polish demonstrate comparable psychometric properties. To that end, we revisit the psychometric properties reported for the American English CAT-CDI (word production) in Kachergis et al. (2022) and compare those to the data from Polish CAT-CDI:Words and Sentences.

<!-- AmE r values copied from Kachergis et al. 2022 -->
```{r}
# set.seed?
cor(by_child_pl$cat_theta, by_child_pl$score_full) # AmE:.86, PL: 0.86
cor(by_child_pl$cat_theta, by_child_pl$full_theta) # AmE: .92, PL: .92
cor(by_child_pl$full_theta, by_child_pl$score_full) # AmE: .95, PL: 0.94

```


We found similarly strong correlations in the two languages between the abilities estimated from CAT-CDI and full CDI scores (English and Polish: $r$ = .86). The correlations were also strong within individual age groups (see Tables \@ref(tab:corr-by-age-eng) and \@ref(tab:corr-by-age-pl)). We also obtained IRT ability estimates from the full CDI using the same 2-PL model. Because these estimates draw on responses to all available items in the full CDI, they incorporate more information about each child’s performance, reducing random error and increasing precision. The 2-PL model further adjusts for item difficulty and discrimination, making these ability scores robust and comparable across individuals. Such ability from the full test version is often considered a baseline for comparisons with ability estimated from a short (adaptive) version. We correlated the abilities estimated from the CAT-CDI and abilities estimated from full CDI and again found strong correlations in both languages, English and Polish: $r$ = .92. Finally, we also correlated the abilities estimated from the full CDI and the full CDI scores, i.e. two measures obtained from the same data. This way we examine consistency between model-based abilities and traditional scores. Again, we found strong correlations in both languages (English: $r$ = .95, Polish: $r$ = `r cor(by_child_pl$full_theta, by_child_pl$score_full)`), confirming that both measures reflect the same underlying ability. 


```{r corr-by-age-eng, include = T}

# TODO:
# 1. merge both tables?
# 2. can we specify place in text where the tables should be placed (rn it seems random)
# ^ See Issue https://github.com/karolina-m/catcdi_validation_eng_pl/issues/5

# English
valid_tab <- prod_s %>% 
  mutate(age_group = cut(age_full, breaks=seq(12,36,3), right=F, include.lowest = T)) %>%
  group_by(age_group) %>%
  summarise(r = cor(fullTheta, catTheta), n=n()) #%>% kableExtra::kable(digits=2)

valid_tab_wide = rbind(round(valid_tab$r, 2), as.character(valid_tab$n))
colnames(valid_tab_wide) = as.character(valid_tab$age_group)
row.names(valid_tab_wide) = c('r ability CAT vs full CDI', 'N')

apa_table(valid_tab_wide, caption="English: Correlations between ability estimated by CAT-CDI and ability estimated from full CDI by children's age.", placement = "H") 
```


```{r corr-by-age-pl, include = T}

age_cor_table <- by_child_pl %>% 
  mutate(age_group = cut(age_CAT, breaks = seq(12, 36, 3), right = FALSE, include.lowest = TRUE)) %>%
  group_by(age_group) %>%
  summarise(r = cor(full_theta, cat_theta), n = n(), .groups = "drop") %>%
  # adding a row for [15,18) age range - present in Kachergis et al. data, absent in Polish
  bind_rows(
    tibble(
      age_group = factor("[15,18)", levels = levels(.$age_group)), # keep factor consistency
      r = NA_real_,
      n = NA_integer_)) %>%
  arrange(age_group)

age_cor_table_wide = rbind(round(age_cor_table$r, 2), as.character(age_cor_table$n))
colnames(age_cor_table_wide) = as.character(age_cor_table$age_group)
row.names(age_cor_table_wide) = c('r ability CAT vs full CDI', 'N')

apa_table(age_cor_table_wide, caption = "Polish: Correlations between ability estimated by CAT-CDI and ability estimated from full CDI by children's age.", placement = "H")
```


### Discrepancy

```{r mse}
# TODO
# 1. Eng values shifted slightly (compared to Kachergis et al. 2022)? 

# Polish
by_child_pl <- by_child_pl %>% mutate(sq_err = (full_theta - cat_theta)^2,
                                      full_cat_diff = full_theta - cat_theta)
mean(by_child_pl$sq_err) # 0.19
median(by_child_pl$sq_err) # 0.08
sd(by_child_pl$sq_err) # 0.45

bad_thresh_pl <- mean(by_child_pl$sq_err) + 1.5*sd(by_child_pl$sq_err)
bad_thresh_pl # 0.86

bad_pl <- subset(by_child_pl, sq_err > bad_thresh_pl)
good_pl <- subset(by_child_pl, sq_err <= bad_thresh_pl) 

by_child_pl <- by_child_pl %>% mutate(extreme_discrep = 
                                           case_when(
                                                sq_err > bad_thresh_pl ~ "yes",
                                                .default = "no"
                                           ))

table(by_child_pl$extreme_discrep)

# English

# mean squared error
prod_s <- prod_s %>% mutate(sq_err = (fullTheta - catTheta)^2,
                            full_cat_diff = fullTheta - catTheta)

bad_thresh_en <- mean(prod_s$sq_err) + 1.5*sd(prod_s$sq_err) # mean=.55, med = .17

bad_en <- subset(prod_s, sq_err > bad_thresh_en) # 18 --> 15?
good_en <- subset(prod_s, sq_err <= bad_thresh_en) # 186 --> 189?

prod_s <- prod_s %>% mutate(extreme_discrep = 
                                           case_when(
                                                sq_err > bad_thresh_en ~ "yes",
                                                .default = "no"
                                           ))

table(prod_s$extreme_discrep)


```

Next, we looked at the mean squared error (MSE) which quantifies the difference between the abilities as estimated by CAT-CDI and the full CDI (for each participant, we subtract the CAT-CDI estimate from the full estimate and square this difference to eliminate negative values). The mean squared error in English was `r mean(prod_s$sq_err)` (*Mdn* = `r median(prod_s$sq_err)`, *SD* = `r sd(prod_s$sq_err)`), and in Polish it was `r mean(by_child_pl$sq_err)` (*Mdn* = `r median(by_child_pl$sq_err)`, *SD* = `r sd(by_child_pl$sq_err)`). The MSE was relatively high in English (compared to Polish), indicating a relatively larger discrepancy between CAT-CDI ability estimates and full CDI ability estimates in English. This occurred despite strong correlations between the two measures. This difference between the full and CAT-CDI estimates may be explained by at least two reasons. First, a number of participants with extreme discrepancies between the CAT-CDI and full CDI estimates may be inflating the MSE. Second, there might be a systematic bias, with one estimate consistently higher than the other. 

```{r}
# re-calculating MSE after excluding extreme cases
prod_s %>% filter(extreme_discrep == "no") %>% summarise(mean(sq_err)) # 0.44, was 0.57
by_child_pl %>% filter(extreme_discrep == "no") %>% summarise(mean(sq_err)) # 0.12, was 0.19
```

To explore the first of these possibilities, we identified children for whom the estimates from the CAT-CDI and full CDI showed extreme discrepancies, i.e. the difference between the two estimates was removed by 1.5 SD from the mean. There were `r sum(prod_s$extreme_discrep == "yes")` such cases (`r round(sum(prod_s$extreme_discrep == "yes")/count(prod_s)*100,2)`%) in the English dataset and `r sum(by_child_pl$extreme_discrep == "yes")` cases (`r round(sum(by_child_pl$extreme_discrep == "yes")/count(prod_s)*100,2)`%) in the Polish dataset. We then re-calculated the mean squared error without these cases, which yielded an MSE of `r round(prod_s %>% filter(extreme_discrep == "no") %>% summarise(mean(sq_err)),2)` (*Mdn* = `r round(prod_s %>% filter(extreme_discrep == "no") %>% summarise(median(sq_err)),2)`, *SD* = `r round(prod_s %>% filter(extreme_discrep == "no") %>% summarise(sd(sq_err)),2)`) in English and `r round(by_child_pl %>% filter(extreme_discrep == "no") %>% summarise(mean(sq_err)),2)` (*Mdn* = `r round(by_child_pl %>% filter(extreme_discrep == "no") %>% summarise(median(sq_err)),2)`, *SD* = `r round(by_child_pl %>% filter(extreme_discrep == "no") %>% summarise(sd(sq_err)),2)`) in Polish. The updated MSE values were only slightly lower than the original ones, suggesting that extreme discrepancies had only a modest effect on the overall error and cannot fully account for the differences between ability estimates from the full CDI and the CAT-CDI in the two languages.

```{r}
# Difference between abilities: full ability minus CAT ability
# Note, this yields mostly negative values (CAT > full) but some positive values as well (CAT < full)

# EN - all
t1 <- apa_print(t.test(prod_s$catTheta, prod_s$fullTheta, paired = T))
t1 # p < 0.001
d1<- cohens_d(prod_s$catTheta, prod_s$fullTheta, paired = T)
d1

# EN - excluding extreme_discrepancy cases (similar)
t.test(good_en$catTheta, good_en$fullTheta, paired = T)
cohens_d(good_en$catTheta, good_en$fullTheta, paired = T)

# PL - all
t2 <- apa_print(t.test(by_child_pl$cat_theta, by_child_pl$full_theta, paired = T)) 
t2 # p < 0.001
d2 <- cohens_d(by_child_pl$cat_theta, by_child_pl$full_theta, paired = T)
d2

# PL - excluding extreme_discrepancy cases (similar)
t.test(good_pl$cat_theta, good_pl$full_theta, paired = T)
cohens_d(good_pl$cat_theta, good_pl$full_theta, paired = T)

```

Next, we examined a possible bias in our two ability estimates. We found that the mean CAT ability estimates were higher than the full CDI ability estimates. That was true both in English, `r t1$full_result` (large effect size: Cohen's $d$ = 1.06, 95% CI $[0.89, 1.23]$), and in Polish, `r t2$full_result` (medium effect size: Cohen's $d$ = 0.73, 95% CI $[0.52, 0.93]$). A closer look showed that the difference between the full and CAT-CDI estimates depended on the child's ability level (see Figure \@ref(fig:diff-fig)). For low ability, the English CAT-CDI estimates tended to be higher than the full scores. For the highest ability, the estimates from the two CDIs seemed most aligned. In Polish, the CAT-CDI estimates generally aligned more closely with the full CDI estimates across the whole range of ability. Still, a similar trend has emerged as in English, though weaker: for the lowest ability levels, the CAT-CDI estimates were mostly higher than the full CDI estimates. However, for the highest ability levels, the CAT-CDI estimates were in fact lower than the full CDI estimates. 


```{r diff-fig, out.width="100%", fig.cap = "The relation between ability (estimated from full CDI) and the size of  discrepancy between the estimates from full CDI and CAT-CDI: (A) English, (B) Polish.", include = T}

# EN
diff_en <- prod_s %>%
  select(subject_id, catTheta, catTheta_SE, fullTheta, fullTheta_SE) %>%
  rename(
    cat_se = catTheta_SE,
    full_se = fullTheta_SE,
    cat_theta = catTheta,
    full_theta = fullTheta) %>%
     mutate(full_theta_bin = cut(full_theta, 
                         breaks = c(-3, -1.5, 0, 1.5, 3.1), 
                         labels = c("[-3,-1.5]", "(-1.5,0]", "(0,1.5]", "(1.5,3]"),
                         include.lowest = TRUE)) %>%
  pivot_longer(
    cols = c(full_theta, full_se, cat_theta, cat_se),
    names_to = c("method", ".value"),
    names_pattern = "(full|cat)_(theta|se)") 

p1 <- diff_en  %>% group_by(full_theta_bin) %>% ggplot(aes(x = full_theta_bin, y = theta, fill = method, color = method)) + geom_boxplot(alpha = 0.5) +
     labs(x = "Theta from Full CDI (binned)", y = "Theta from each CDI version", title = "A: English") +
     theme_light() + 
     theme(legend.position = "bottom", legend.title = element_blank()) +
     ylim(-3.5,3.5)


# PL

diff_pl <- by_child_pl %>%
  select(child_id, cat_theta, cat_theta_se, full_theta, full_theta_se) %>%
  rename(
    cat_se = cat_theta_se,
    full_se = full_theta_se) %>%
     mutate(full_theta_bin = cut(full_theta, 
                         breaks = c(-3, -1.5, 0, 1.5, 3), 
                         labels = c("[-3,-1.5]", "(-1.5,0]", "(0,1.5]", "(1.5,3]"),
                         include.lowest = TRUE)) %>%
  pivot_longer(
    cols = c(full_theta, full_se, cat_theta, cat_se),
    names_to = c("method", ".value"),
    names_pattern = "(full|cat)_(theta|se)")

p2 <- diff_pl  %>% group_by(full_theta_bin) %>% ggplot(aes(x = full_theta_bin, y = theta, fill = method, color = method)) + geom_boxplot(alpha = 0.5) +
     labs(x = "Theta from Full CDI (binned)", y = "Theta from each CDI version", title = "B: Polish") +
     theme_light() + 
     theme(legend.position = "bottom", legend.title = element_blank()) +
     ylim(-3.5,3.5)


fig_theta <- ggarrange(p1, p2, ncol = 2, common.legend = T, legend = "bottom")
fig_theta
ggsave("fig_theta.jpeg", dpi = 300, width = 200, height = 150, units = "mm")



```

The question that remains is whether the differences between the two estimates of abilities come from CAT over-estimating child's true lexical ability or from the full CDI underestimating child's ability. For one, it may be that in CAT-CDI parents answer "yes, my child produces this word" to more items that would be expected (Kachergis et al. 2022), and as a result, CAT-CDI may overestimate the child's true ability. On the other hand, it could be that as the full CDI is very long, parents may omit marking some items that their child produces (see e.g., Łuniewska et al., 2025 for parental inconsistencies in reporting vocabulary of older children). This could result in the full CDI underestimating the child's true ability. It is possible that both are true and the question is difficult to answer without direct assessment of child's lexical ability. With the existing data, we can at least compare the precision of the two estimates to assess which measure is likely more reliable.


## Precision of the estimates

We looked at how precision of the estimate, measured by standard error of the theta (SE), related to the ability levels. In both languages, the characteristic U-shaped curves indicate that measurement precision is the highest around the middle of the ability range (i.e., showing lowest SE) and decline at the extremes (very low or very high ability the SE values are the highest) (see Figure \@ref(fig:fig-se-by-theta)). This U-shape is observed for both the CAT-CDI and the full CDI, indicating that even the full CDI provides relatively less precise estimates at the extreme ends of ability. Note that in Polish (Figure \@ref(fig:fig-se-by-theta), panel B) the lowest ability estimates (i.e., [-3,-1.5]) yielded a slightly lower median SE in CAT-CDI than in the full CDI, suggesting that at lowest ability levels, CAT-CDI might be similarly or slightly more precise than the full CDI. For the highest ability estimates (i.e., (1.5, 3]) the full CDI came with lower median SE than the CAT-CDI, even in many cases, falling below the SE of 0.2 (a cut-off suggested by Makransky et al., XXX).

```{r fig-se-by-theta, out.width="100%", fig.cap = "The relation between ability estimaes from full CDI and CAT-CDI and the standard error of that estimate: (A) English, (B) Polish.", include = T}

# EN
se_en <- prod_s %>%
  select(subject_id, catTheta, catTheta_SE, fullTheta, fullTheta_SE) %>%
  rename(
    cat_se = catTheta_SE,
    full_se = fullTheta_SE,
    cat_theta = catTheta,
    full_theta = fullTheta) %>%
  pivot_longer(
    cols = c(full_theta, full_se, cat_theta, cat_se),
    names_to = c("method", ".value"),
    names_pattern = "(full|cat)_(theta|se)") %>%
     mutate(theta_bin = cut(theta, 
                         breaks = c(-3, -1.5, 0, 1.5, 3.1), 
                         labels = c("[-3,-1.5]", "(-1.5,0]", "(0,1.5]", "(1.5,3]"),
                         include.lowest = TRUE))

p3_alt <- se_en %>%
  ggplot(aes(x = theta, y = se, color = method, fill = method)) + 
  geom_smooth(se = FALSE, method = "loess", span = 0.5, size = 1.2) +
  geom_boxplot(
  aes(group = interaction(theta_bin, method)), 
  alpha = 0.2, size = 0.1,
  width = 0.3) +   # fixed width, all equal
  scale_x_continuous(
    breaks = c(-2.5, -1, 1, 2.5),
    labels = c("[-3,-1.5]", "(-1.5,0]", "(0,1.5]", "(1.5,3]")) +
labs(y = "SE", x = "Theta", title = "A: ENGLISH") +
     ylim(0, 0.5) +
  theme_light() + theme(legend.title = element_blank())


# PL
se_pl <- by_child_pl %>%
  select(child_id, cat_theta, cat_theta_se, full_theta, full_theta_se) %>%
  rename(
    cat_se = cat_theta_se,
    full_se = full_theta_se) %>%
  pivot_longer(
    cols = c(full_theta, full_se, cat_theta, cat_se),
    names_to = c("method", ".value"),
    names_pattern = "(full|cat)_(theta|se)") %>%
     mutate(theta_bin = cut(theta, breaks = seq(-3, 3, by = 1.5), include.lowest = TRUE))

p4_alt <- se_pl %>%
  ggplot(aes(x = theta, y = se, color = method, fill = method)) + 
  geom_smooth(se = FALSE, method = "loess", span = 0.5, size = 1.2) +
  geom_boxplot(
  aes(group = interaction(theta_bin, method)), 
  alpha = 0.2, size = 0.1,
  width = 0.3) +   # fixed width, all equal
  scale_x_continuous(
    breaks = c(-2.5, -1, 1, 2.5),
    labels = c("[-3,-1.5]", "(-1.5,0]", "(0,1.5]", "(1.5,3]")) +
labs(y = "SE", x = "Theta", title = "B: POLISH") +
     ylim(0, 0.5) +
  theme_light() + theme(legend.title = element_blank())

fig_se_by_theta_alt <- ggarrange(p3_alt, p4_alt, ncol = 2, common.legend = T, legend = "bottom") %>%
     annotate_figure(fig.lab = NULL,
                     bottom = text_grob("Note that in English there is only one case where CAT estimate was smaller than -1.5,\n therefore no CAT boxplot is plotted in that bin.", hjust = 0, x = 0, face = "italic", size = 10))
     
fig_se_by_theta_alt
ggsave("fig_se_by_theta_alt.jpeg", dpi = 300, width = 200, height = 150, units = "mm")


```



In both CAT-CDIs, we have been able to reliably estimate lexical ability for the majority of children. Taking as the cut-off Makransky's suggestion to consider an SE below 0.2 as acceptable, such precision of the CAT-CDI estimate was obtained for `r sum(prod_s$catTheta_SE < 0.2)` out of `r length(prod_s$subject_id)` American participants (`r round(sum(prod_s$catTheta_SE < 0.2)/count(prod_s)*100,1)`%) and `r sum(by_child_pl$cat_theta_se < 0.2)` out of `r length(by_child_pl$child_id)` Polish participants (`r round(sum(by_child_pl$cat_theta_se < 0.2)/count(by_child_pl)*100,1)`%). The remaining participants were at the extremes of the ability scale (see Figure \@ref(fig:se-by-theta-fig)). The ability estimates from the Polish CAT-CDI generally fell well below an SE of 0.2, except for the highest-ability observations. Because the Polish CAT-CDI was designed to stop once an SE of 0.1 was reached, the precision of Polish estimates never went below 0.1 (i.e., the test was stopped at this point). 

Taken together, the results reported so far show that for the English CAT-CDI, the CAT-CDI estimates correlate strongly with both full CDI scores and estimates obtained from the full CDI. Importantly, for the large majority of children, the obtained CAT-CDI estimates were below the precision cut-off point of 0.2. For the lowest abilities, the CAT-CDI estimates are higher than those of the full CDI, but in the Polish data, the precision of the CAT-CDI estimates is slightly higher than that of the full CDI. This might potentially suggest that for lower abilities, CAT estimates may be more aligned with the true abilities, however, more research needs to be done.

### Old

```{r}
#  SEM

# PL
mean(by_child_pl$cat_theta_se) # 0.11
median(by_child_pl$cat_theta_se) # 0.10

# EN
mean(prod_s$catTheta_SE) # 0.17
median(prod_s$catTheta_SE) # 0.16

t2 <- apa_print(t.test(prod_s$catTheta_SE, by_child_pl$cat_theta_se)) # p < 0.001

cor_age <- apa_print(cor.test(prod_s$catTheta_SE, prod_s$age_full)) # watch out, SE is weakly but significantly correlated with age, and the English validation sample is slightly (but significantly) younger than the Polish one. This differece in SEM between the samples can be partially pulled by age.
```

The two CAT-CDIs differed in their settings as to when to finish the administration. Specifically, the Polish CAT-CDI was set so that the SEM as low as 0.1 be reached (with as few items as possible), and if the SEM could not be lowered to or below 0.1, the CAT-CDI stopped after administering a maximum of 75 items. The English CAT-CDI on the other hand was set to administer a maximum of 50 items, but the administration could be stopped earlier if the SEM of 0.15 or lower was reached. In other words, the Polish CAT-CDI prioritized as low SEM as possible, while the English CAT-CDI prioritized shorter test length. This difference resulted in slightly lower mean SEM in Polish CAT-CDI, compared to English CAT-CDI `r t2$full_result` but comparable mean number of items administered, `r t1$full_result`. 

```{r}
# SEM percentiles

# PL
ecdf_fun <- ecdf(as.numeric(by_child_pl$cat_theta_se))
by_child_pl <- by_child_pl %>%
  mutate(cat_theta_se_percentile = ecdf_fun(cat_theta_se) * 100) %>%
     relocate(cat_theta_se_percentile, .after = cat_theta_se)

# EN
ecdf_fun <- ecdf(as.numeric(prod_s$catTheta_SE))
prod_s <- prod_s %>%
  mutate(catTheta_SE_percentile = ecdf_fun(catTheta_SE) * 100) %>%
     relocate(catTheta_SE_percentile, .after = catTheta_SE)

round(sum(by_child_pl$cat_theta_se < 0.2)/count(by_child_pl)*100,1) # 96.5%
round(sum(prod_s$catTheta_SE < 0.2)/count(prod_s)*100,1) # 92.6%

```



## Parental consistency

```{r}
# PL
consistency_pl_sum %>% ungroup()%>% dplyr::summarise(mean = mean(perc, na.rm = T),
                                    median = median(perc, na.rm = T),
                                    max = max(perc, na.rm = T),
                                    min = min(perc, na.rm = T))

# EN
consistency_en_sum %>% ungroup()%>% dplyr::summarise(mean = mean(perc, na.rm = T),
                                    median = median(perc, na.rm = T),
                                    max = max(perc, na.rm = T),
                                    min = min(perc, na.rm = T))

# MERGE
consistency_en_sum$lang <- "ENGLISH"
consistency_pl_sum$lang <- "POLISH"

consistency_en_sum <- consistency_en_sum %>% rename(id = subject_id)

consistency <- consistency_en_sum %>% select(id, perc, lang)
consistency_pl_sum %>% select(id, perc, lang) %>% rbind(consistency) -> consistency
consistency <- consistency %>% mutate(perc = perc*100)
```

```{r parental-consistency-perc-fig, fig.width = 10, fig.height = 5, fig.cap = "Percentage of parents' consistency in responding to items in the two CDI versions in the same way (in the English and in the Polish dataset).", fig.pos = '!h', include = T}
consistency %>% ggplot(aes(x = lang, y = perc)) + 
     geom_boxplot(width = 0.5) + geom_jitter(alpha = 0.3, width = 0.3) +
     labs(y = "Parental consistency in %", x = "") + theme_light() + ylim(0,100)
```


We compared the parental consistency in their responses in the full and CAT version of CDI in two ways - using a percentage of agreement and by calculating Cohen's Kappa, which additionally factors in chance agreement. When calculating the percentage agreement, for each parent we calculated the number of items to which they responded in the same way in both CDI versions against the whole number of items asked in both CDI versions. We found though the Median consistency was relatively high: `r round((median(consistency_en_sum$perc, na.rm = T)*100),2)`% for English and `r round((median(consistency_pl_sum$perc, na.rm = T)*100),2)`% for Polish, it varied largely across parents in both samples (see Figure \@ref(fig:parental-consistency-perc-fig)). 

```{r}
# Cohen's kappa
# An overall Cohen's kappa wouldn't work here, because each parent saw a (partially) different subset of items in their CAT administration. So what I calculate a Cohen's kappa for each parent and then I take an average/median across the participants.
# I'm open to ideas: https://github.com/karolina-m/catcdi_validation_eng_pl/issues/6 

library(irr)

# PL
kappa_pl <- consistency_pl %>% select(id, definition, answer_full, answer_cat)

# Function to compute kappa for one parent
compute_kappa <- function(data) {
  if (nrow(data) < 5) {
    return(NA)  # Too few items — return NA or percent agreement instead
  } else {
    ratings <- data %>%
      select(answer_full, answer_cat)
    return(kappa2(ratings)$value)
  }
}

# Compute kappa per parent
kappa_results_pl <- kappa_pl %>%
  group_by(id) %>%
  group_modify(~ tibble(kappa = compute_kappa(.x))) %>%
  ungroup()

round(mean(kappa_results_pl$kappa, na.rm = T),2) #0.39
round(median(kappa_results_pl$kappa, na.rm = T),2) #0.38

# EN
kappa_en <- consistency_en %>% select(subject_id, definition, response_full, response_cat)

# Function to compute kappa for one parent
compute_kappa <- function(data) {
  if (nrow(data) < 5) {
    return(NA)  # Too few items — return NA or percent agreement instead
  } else {
    ratings <- data %>%
      select(response_full, response_cat)
    return(kappa2(ratings)$value)
  }
}

# Compute kappa per parent
kappa_results_en <- kappa_en %>%
  group_by(subject_id) %>%
  group_modify(~ tibble(kappa = compute_kappa(.x))) %>%
  ungroup()

# en
round(mean(kappa_results_en$kappa, na.rm = T),2) #0.4
round(median(kappa_results_en$kappa, na.rm = T),2) #0.39


```


We also calculated Cohen's kappa per each parent, considering their observed agreement (between responses to the same items in full and CAT CDI) and expected agreement by chance. The kappa values indicated fair (or close to fair) agreement: English *M* =  `r round(mean(kappa_results_en$kappa, na.rm = T),2) #0.4`, *Mdn* = `r round(median(kappa_results_en$kappa, na.rm = T),2)`, Polish *M* =  `r round(mean(kappa_results_pl$kappa, na.rm = T),2) #0.4`, *Mdn* = `r round(median(kappa_results_pl$kappa, na.rm = T),2)`. *NOTE*: [There's an Issue for that here](https://github.com/karolina-m/catcdi_validation_eng_pl/issues/6)

## Item selection in the two CAT-CDIs


```{r}
# Data prep

# POLISH
items_run_pl <- responses_cat_pl %>% select(definition) %>% group_by(definition) %>% summarise(n = length(definition))
items_run_pl <- merge(items_cat_pl, items_run_pl, by.x = "item", by.y = "definition", all.x = T)
items_run_pl[is.na(items_run_pl)] <- 0
items_run_pl$never <- ifelse(items_run_pl$n == 0, "never", "at_least_once")

items_run_pl <- items_run_pl %>% 
  mutate(items_perc = (n/115)*100) %>%
  mutate(percent_for_color = ifelse(items_perc == 0, NA, items_perc)) 

# AMERICAN ENGLISH

items_run_en <- resps %>%
  select(subject_id, response_cat, definition)

items_run_en <- items_run_en %>% mutate(
  item_used = case_when(
    response_cat == "no_test" ~ 0,
    .default = 1
  )
)

items_run_en <- items_run_en %>%
  select(definition, item_used) %>%
  group_by(definition) %>%
  mutate(n = sum(item_used)) %>%
  select(definition, n) %>% unique()

# how many participants in resps?
resps %>% select(subject_id) %>% unique() %>% nrow()

items_run_en <- items_run_en %>% 
  mutate(items_perc = (n/204)*100) %>%
  mutate(percent_for_color = ifelse(items_perc == 0, NA, items_perc)) 

items_run_en <- merge(items_run_en, coefs_2pl[,c("definition", "a1", "d")], 
                   all.x = T,
                   by = "definition")

# ame items lack categories
eng_ws_items <- read_csv("data/english/eng_ws_items.csv")
items_run_en <- merge(items_run_en, eng_ws_items[,c("definition", "category")], all.x = T, by = "definition")
rm(eng_ws_items)
```

```{r}
# summary
# PL
sum(items_run_pl$n != 0) # 258
round(sum(items_run_pl$n != 0)/count(items_run_pl)*100,2) # 38.74

# EN
sum(items_run_en$n != 0) # 251
round(sum(items_run_en$n != 0)/count(items_run_en)*100,2) # 36.91
```

```{r}
# low discrimination into seperate df
low_disc <- items_run_en %>% filter(a1 <2 & d > -1.5 & d < 1.5)
```

It is to be expected that a CAT-CDI will not need to administer all the items in the item bank. In fact, in all the task administrations in the validation study the English CAT-CDI used `r sum(items_run_en$n != 0)` items (`r round(sum(items_run_en$n != 0)/count(items_run_en)*100,2)`%) and the Polish CAT-CDI used `r sum(items_run_pl$n != 0)` items (`r round(sum(items_run_pl$n != 0)/count(items_run_pl)*100,2)`%).
By design, a CAT-CDI selects the items that provide the most information for each participant. In practice, this means it draws from a subset of items that both match the participant’s current ability estimate in difficulty and show high discrimination, allowing clear distinctions between individuals with similar ability levels. A potential concern, however, is that the CAT-CDI might repeatedly rely on the same subset of optimal items. CAT-CDI's usefulness in research—particularly in longitudinal studies—lies in the possibility of administering the CDI frequently without parents becoming overly familiar with specific, frequently repeated items. This advantage holds only if the CAT-CDI does not consistently present the same items across administrations. Therefore, we investigated whether the two CAT-CDIs repeatedly selected any items too frequently. In English, 3 items appeared in more than half of administrations - *"long"* in 64% of administrations, *"make"* in 54% of administrations, and *"last"* in 53% of administrations. As expected, they are also high discrimination items in their difficulty bins. "Long" is additionally one of the starting items, i.e. a first item shown to the parent, chosen so that there is high probability the parent can respond positively. In Polish, 3 items appeared in more than half of administrations - *"szukać"* (to look for) appearing in 83% of administrations, *"znaleźć"* (to find) appearing in 61% of administrations, and *"babcia"* (grandmother) appearing in 60% of administrations. As expected, all three items show very high discrimination. and *"Szukać"* (to look for) and *"znaleźć"* (to find) are in fact in top 5 discrimination items in general. *"Babcia"* (grandmother) is one of the starting items. The fact that only 3 items out of c.a. 250 (c.a. 1%) are used in more than half administrations is to CAT-CDI's advantage. 

```{r}
items_en_pl <- merge(items_cat_pl, items_cat_en, 
                  all.x = T, all.y = T,
                  suffixes = c("_pl","_en"),
                  by = "uni_lemma") %>% filter(!is.na(item) & !is.na(definition))
count(items_en_pl) # 390

items_en_pl <- merge(items_en_pl, items_run_pl[,c("item", "items_perc")], all.x = T, by = "item")
names(items_en_pl)[names(items_en_pl) == "items_perc"] <- "items_perc_pl"

items_en_pl <- merge(items_en_pl, items_run_en[,c("definition", "items_perc")], all.x = T, by = "definition")
names(items_en_pl)[names(items_en_pl) == "items_perc"] <- "items_perc_en"

c3 <- apa_print(cor.test(items_en_pl$items_perc_pl, items_en_pl$items_perc_en, paired = T, method = "spearman")) # .37

```


To check whether the CDI items common in both languages may show similar frequencies in CAT-CDI administrations, we ran Spearman's rank correlation and found that the items frequencies in administrations were weakly correlated, `r c3$full_result`. However, the two CAT-CDIs differed in their potential length, as the English CAT-CDI was set to administer 25--50 items and the Polish CAT-CDI could administer up to 75 items (no minimum was set, but in practice the minimum items administered were 13). As an item is more likely to appear in longer tests (simply due to test length), then frequency of the Polish items could be inflated. *NOTE*:Could it be that, as an item is more likely to appear in longer tests (simply due to test length), frequency of the Polish items could be inflated? Could it influence the item's rank? I could probably recalculate item frequency not per total number of administrations, but per total number of items administered? [SEE ISSUE #3](https://github.com/karolina-m/catcdi_validation_eng_pl/issues/3)

```{r}

```


## Item properties in the two CAT-CDIs
Our second aim was to analyze similarities and differences in IRT item properties in CAT in English and Polish.

There are 679 items in the English CAT-CDI and 666 items in the Polish CAT-CDI. In both languages, the items' difficulty and discrimination parameters were calculated using IRT 2 parameter model (these included separate samples, see Kachergis et al. 2022 and Krajewski et al., in preparation). An item’s difficulty indicates the ability level at which there is a 50% probability that a participant will respond correctly (here: that a child will know a word). Difficulty is expressed on the same scale as theta (measured ability): negative values correspond to harder items, values near zero to items of moderate difficulty, and positive values to easier items (note that difficulty is inversely related to theta and so, negative theta means low ability while positive theta means high ability). An item’s discrimination shows how well it can distinguish between individuals whose abilities are close. In this paper, we focus mainly on item difficulty, as it is directly tied to the ability being measured, whereas discrimination reflects more the quality of the item as a measurement tool.

```{r diff_across_lgs}
# item difficulty by language 
items_cat_en %>% 
     dplyr::summarise(min(d), max(d), mean(d), median(d), sd(d))

items_cat_pl %>% 
     dplyr::summarise(min(d), max(d), mean(d), median(d), sd(d))

t1 <- apa_print(t.test(items_cat_en$d, items_cat_pl$d))

```

```{r common_items}
# common items - still different mean difficulty?
t2 <- apa_print(t.test(items_en_pl$d_en, items_en_pl$d_pl, paired = T))
# t = -23.112, df = 389, p-value < 2.2e-16

# visualise
items_en_pl %>% select(uni_lemma, d_pl, d_en) %>% pivot_longer(cols = d_pl:d_en, names_to = "lang", values_to = "d") %>% ggplot(aes(x = lang, y = d, group = lang)) + geom_boxplot() + geom_jitter(alpha = 0.3)

items_en_pl %>% select(uni_lemma, d_pl, d_en) %>% 
     ggplot(aes(x = d_pl, y = d_en)) + geom_jitter(alpha = 0.3) + geom_smooth(method = "lm") +
     stat_cor(cor.coef.name = "r", p.accuracy = 0.001, r.accuracy = 0.01, size = 5, show.legend = F) +
     theme_light() + labs(x = "Item difficulty in Polish", y = "Item difficulty in English")

# corr
c <- apa_print(cor.test(items_en_pl$d_pl, items_en_pl$d_en)) # 0.65
```

English items were in general more difficult than the Polish items, `r t1$full_result` (English: min = `r round(min(items_cat_en$d),2)`, max = `r round(max(items_cat_en$d),2)`, *M* = `r round(mean(items_cat_en$d),2)`, *Mdn* = `r round(median(items_cat_en$d),2)`, *SD* = `r round(sd(items_cat_en$d),2)`; Polish: min = `r round(min(items_cat_pl$d),2)`, max = `r round(max(items_cat_pl$d),2)`, *M* = `r round(mean(items_cat_pl$d),2)`, *Mdn* = `r round(median(items_cat_pl$d),2)`, *SD* = `r round(sd(items_cat_pl$d),2)`). This was true also for a subset of `r count(items_en_pl)` items common to both languages - English items still proved to be more difficult than the Polish ones: `r t2$full_result`. This is a result of the characteristics of the samples used to estimate the item parameters. In English, item difficulty was calculated based on a broader sample of children aged 12–36 months, spanning the CDI:WG, CDI:WS, and CDI-III (reflecting the idea of having one CAT-CDI for all the age versions), whereas the Polish data came from a sample of narrower age range of 18–36 months, corresponding to the CDI:WS. As a result, item difficulty in English was estimated using a relatively younger sample, for whom certain items may have been more challenging -- thus appearing more difficult -- compared to the older Polish sample. Still the item difficulty for common items in the two languages was positively and moderately correlated: `r c$full_result`.

We also examined whether CDI semantic categories show comparable difficulty values across the two languages. To this end, we conducted a series of Spearman’s rank correlations on item difficulty within each semantic category (we considered only items common to both the Polish and English CDIs) (see Table \@ref(tab:d-by-cdi-category-tab)). The rank correlations coefficients varied by category, but for half of the categories the correlations were moderate to strong (0.52 to 0.91). 

The least correlated semantic category was Quantifiers. The discrepancy may be partially to the fact that many English quantifiers map onto multiple Polish equivalents - for example, "some" maps to two Polish CDI items, "jeszcze" (with a similar difficulty value as the English "some") and "więcej" (more difficult than "jeszcze" and "some"). These two Polish items are used in different contexts, i.e., "jeszcze" would be used more often in child directed speech than "więcej" also because it can be used to refer to both quantity and repetitive actions (meaning "again"). "Więcej" implies a quantitative comparison of some sorts, as in "more than before" or "more than two".

Interestingly, the second least correlated semantic category were Clothes. Here, some divergence may be explained by one-to-multiple mapping (e.g., "hat" is either "czapka" or "kapelusz" in Polish), but also words characteristics, e.g., "bib" in Polish is a much more complex word, "śliniaczek"; English "underpants" are more morphologically complex (and maybe less frequent) than Polish "majtki".

Taken together, these two semantic categories reflect the existing cross-linguistic differences in lexical mapping as well as item-specific factors such as frequency, morphological complexity, and contextual usage that .

```{r d-by-cdi-category-tab, include = T}
# correlations by cdi semantic category on common words
# spearman because ranked

items_en_pl %>% 
     mutate(category_pl = 
                 fct_recode(category_pl,
                    "descriptive_words" = "descriptive_words (adverbs)")) %>%
     mutate(category_pl = 
                 fct_recode(category_pl,
                    "descriptive_words" = "descriptive_words (adjectives)")) %>%
     group_by(category_pl) %>% summarise(
     rho = cor(d_pl, d_en, method = "spearman", use = "complete.obs"),
     n = length(category_pl)) %>% arrange(rho) %>%
     rename("CDI category" = "category_pl") -> corr_by_category

apa_table(corr_by_category, caption="Cross-linguistic correlations of item difficulty within each CDI semantic category (Spearman's rank correlations).", placement = "H")
```


Last, we investigated whether particular lexical categories show related difficulty values across the two languages. We considered the following lexical classes: verbs, nouns, adjectives, function words and other, i.e. mostly sounds and words relating to routines (e.g. "hello", "yes", "thank you") and time (e.g. "tommorow", "day"). Again, we performed a series of Spearman's rank correlations (on the difficulty of items common to CDI-CATs in Polish and English) for each lexical class (see Table \@ref(tab:d-by-lexical-class-tab). The weakest correlation was for function words (0.40), then verbs (0.47), adjectives (0.53), nouns (0.6), other (0.75). Notably, function words include many of the CDI semantic categories which showed the lowest correlations in Table \@ref(tab:d-by-cdi-category-tab), i.e., quantifiers, connecting words, locations, and pronouns. Again, many English function words map onto multiple Polish equivalents, e.g., "which" corresponds to "jaki" (when asking about type, kind, or quality) and "który" (when selecting from a limited set of options). These cross-linguistic differences partially explain the weaker correlation for function words. What is more, because function words are structurally tied to grammar (which is highly language-specific) they naturally vary in frequency in child-directed speech and in their communicative salience, which can affect how well children know them across languages. By contrast, the relatively high correlations for nouns, sounds and words relating to routines (grouped together as "other") reflect their consistent easiness/difficulty across languages: words such as "mum", "ball", "hi", "woof woof" are acquired early, aligning with findings that nouns are strongly and consistently over-represented in early vocabularies across many languages (e.g., Frank et al., 2021). 

```{r d-by-lexical-class-tab, include = T}
# correlations by lexical class on common words
# spearman because ranked

# PL
items_cat_pl <- merge(items_cat_pl, items_full_pl[,c("item_id", "lexical_class")], all.x = T, 
                      by.x = "itemID", by.y = "item_id") 

items_full_en <- read.csv("data/english/eng_ws_items.csv")
items_cat_en <- merge(items_cat_en, items_full_en[,c("definition", "lexical_class")], all.x = T, 
                      by = "definition") 


items_en_pl <- merge(items_cat_pl, items_cat_en, 
                  all.x = T, all.y = T,
                  suffixes = c("_pl","_en"),
                  by = "uni_lemma") %>% filter(!is.na(item) & !is.na(definition))

items_en_pl %>% group_by(lexical_class_en) %>% summarise(
     rho = cor(d_pl, d_en, method = "spearman", use = "complete.obs"),
     n = length(lexical_class_en)) %>% arrange(rho) -> corr_by_lex_class

apa_table(corr_by_lex_class, placement = "H")

# add d_diff = differences in item difficulty between PL and ENG:
items_en_pl <- items_en_pl %>% mutate(d_diff = abs(d_en - d_pl))

```



# Discussion


It seems then that the relation between the two estimates differs, depending on the ability level: for lower abilities, the difference between the two estimates is most visible, with CAT-CDI estimates higher than the full CDI estimtes. For higher ability, the difference decreases or even (as in the case of Polish) changes direction, with CAT-CDI estimates lower than the full CDI estimates.

In points for now:    

1. *Strong correlations with full CDI*.   
The CAT-CDIs were strongly correlated with the full CDI versions in both languages, which supports the overall validity of the CAT approach. These correlations held for both monolingual and multilingual participants. However, the Polish sample was small, so further research is needed—and planned—to confirm these findings.    

2. *Cases of extreme divergence*.   
CAT-CDI scores diverged notably from full CDI estimates for only a small number of children in both language groups. In the Polish sample all divergent cases had low SEM, and their full CDIs were unusually short--maybe the full form underestimated their ability? In the English sample, their full CDIs were not very short, but SEM was relatively low, which is good.     

3. *Item paramteres across languages*.   
Item parameters (i.e., difficulty and discrimination) showed moderate correlations across languages. These likely stem from variations in the calibration samples. Still, some semantic categories showed consistent patterns across languages, suggesting underlying structural similarities in lexical development.   

4. *Item selection*.   
Each CAT version used c.a. half of the full CDI item pool. Few items were overused (mostly those with highest discrimination in the item pool or starting items). But we also saw items being chosen for CAT that wouldn't be considered very informative (i.e., of low discrimination). That's something to explore further, I wrote to Phil Chalmers (author of mirt and mirtCAT), maybe he'll be able to help?    

5. *Efficiency and Reliability*.   
CAT-CDIs were significantly shorter than the full CDIs while maintaining high reliability for most children. The few less reliable estimates were for children at the ends of the ability range (children with either very low or very high lexical skills). Comparing Polish and English samples, the ability ranges with low SEM differed (see Figure 3) - this again might be a result of the calibration samples and item parameters? This is a good place to stress the importance of well-balanced, diverse calibration datasets. 
Also, while the CAT-CDI is efficient and useful for screening or initial identification (“pomiar przesiewowy” in Polish), it may be less suited for diagnostic purposes where high precision is needed across the full ability range.

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::

`r cite_r()`
