---
title             : ""
shorttitle        : "Supplementary Material"
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}

library(papaja)
library(tidyverse)
library(Multilada)
library(here)
library(effectsize)

```

```{r analysis-preferences}
# Seed for random number generation
set.seed(123)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = F, include = F, warning = F, message = F, out.width="100%")
```


```{r, echo = F}
# Data
source(here("data/polish/data_polish.R"))
```

The Polish validation study included `r sum(by_child_pl$lang_group == "multilingual")` data from bi- and multilingual families. Though it is a small group, we decided to explore their correlation coefficients (non-parametric Spearman's rho) and found these were similar to those found for Polish monolingual children (see Table \@ref(tab:corr-by-langgroup) in Supplementary Materials). *NOTE*: [Here is an Issue for commenting on Supplementary Materials](https://github.com/karolina-m/catcdi_validation_eng_pl/issues/5). Since the correlation coefficients were similar in both groups, and as the bi- and multilingual group was was too small to warrant a separate analysis, we decided to include them in the Polish dataset alongside the Polish monolinguals.

```{r corr-by-langgroup, include = T}

# TODO: 
# 1. the table could use some re-styling (lgs as columns?)
# 2. move to Supplementary (separate .Rmd?)

cor1 <- by_child_pl %>%
  group_by(lang_group) %>%
  summarise(r = cor(cat_theta, score_full, method = "spearman"), n = n()) %>%
  mutate(correlation = "Ability from CDI-CAT ~ full CDI score")

cor2 <- by_child_pl %>%
  group_by(lang_group) %>%
  summarise(r = cor(cat_theta, full_theta, method = "spearman"), n = n()) %>%
  mutate(correlation = "Ability from CDI-CAT ~ ability from full CDI")

cor3 <- by_child_pl %>%
  group_by(lang_group) %>%
  summarise(r = cor(full_theta, score_full, method = "spearman"), n = n()) %>%
  mutate(correlation = "Ability from full CDI ~ full CDI score")

lang_cor_table <- bind_rows(cor1, cor2, cor3)
rm(cor1, cor2, cor3)

apa_table(lang_cor_table, caption = "Table S1 - Spearman's correlations for monolingual and multilingual children in the Polish dataset", placement = "H")

```
